---
title: "Google Spreadsheets connection instructions"
format: gfm
editor: source
execute: 
  eval: false
---

```{r setup}
# Set working directory
knitr::opts_knit$set(root.dir = here::here())
```

# Introduction

This document explains how to authorize and configure a connection to Google
Spreadsheets, so it can be used as a back-end to store the application data.

The procedure consists of setting up and encrypted [OAuth
token](https://oauth.net/) for granting the Shiny app access to Google
Spreadsheets. After setting it up, the token can be read and used in the app for
gaining access to the Spreadsheets API through [the {googlesheets4} R
package](https://googlesheets4.tidyverse.org/).

## Attributions

This procedure uses [the {gargle} package](https://gargle.r-lib.org/) (an R
interface for the Google APIs) to authenticate. Its description is based on the
following sources, both of them articles in the package website:

-   [Non-interactive
    auth](https://gargle.r-lib.org/articles/non-interactive-auth.html#project-level-oauth-cache),
    specially [section "Project-level OAuth
    cache"](https://gargle.r-lib.org/articles/non-interactive-auth.html#project-level-oauth-cache)

-   [Managing tokens
    securely](https://gargle.r-lib.org/articles/managing-tokens-securely.html),
    specially [section "Encrypt an R
    object"](https://gargle.r-lib.org/articles/managing-tokens-securely.html#encrypt-an-r-object)

## Setup summary

The setup process consists of the following steps:

1.  Set up an OAuth token

2.  Create an encryption key

3.  Set the encryption key as an environment variable

4.  Encrypt the OAuth token

# Detailed procedure

## Set up an OAuth token

This step will create an OAuth token for the Shiny app to use non-interactively,
which is necessary to deploy it on [shinyapps.io](https://www.shinyapps.io/)
without human intervention.

We will make use of the interactive capabilities of {gargle} to generate a token
through a browser and cache it in a secret directory (i.e. folder), local to
this project.

### Procedure

1.  Designate a project-specific cache directory to be used by {gargle}

```{r setup-cache-dir}
CACHE_PATH <- ".secrets"
options(gargle_oauth_cache = CACHE_PATH)
```

2.  Purposely trigger interactive authorization by {googlesheets4}

```{r interactive-auth}
googlesheets4::gs4_auth()
```

3.  In the browser tab opened by the call to `googlesheets4::gs4_auth()`, click
    on your account of choice to "authorize *Tidyverse API Packages*".

4.  In the next page, click on the button "Continue"

5.  Check the box "View, edit, create and delete all you spreadsheets in Google
    Spreadsheets

6.  Click on the button "Continue"

7.  If the authorization has been sucessfull, a page will show the following
    message:

```         
Authentication complete. Please close this page and return to R.
```

Close this window/tab and return to Rstudio. Now {googlesheets4} can read the
token from the designated cache:

```{r read-token}
gs_token <- googlesheets4::gs4_token()
```

{googlesheets4} will use this token as long as it finds it in the cache folder,
without the need for interactively authenticating again.

## Create an encryption key

After completing the previous step, you will notice that a ".secrets" folder has
been created in the project root folder, which contains a file with a name
similar to "<token_id>\_<user_name>@gmail.com", where `<token_id>` is a
hexadecimal number with several digits, and `<user_name>` is the user name of
the Google account used to "authorize *Tidyverse API Packages*".

This file is the token that will be used to grant the Shiny app access to the
spreadsheet file. However, as this file needs to be deployed to
[shinyapps.io](https://www.shinyapps.io/) along with the app code and associated
files, it needs to be encrypted to prevent it from the risk of being exposed
publicly. This could allow malicious agents accessing not only the spreadsheet
for this project, but all other spreadsheets this account has access to.

Instead, the token is encrypted with a symmetric key (i.e. a "password") that
will be used in the app to decrypt it.

### Procedure

1.  Generate a secret key using {gargle}

```{r gen-key}
secret_key <- gargle::secret_make_key()
```

## Set the encryption key as an environment variable

Using the encryption key requires that it is set as an environment variable.
This prevents the key from being used by calling its value directly, which could
result in logging its value in log files that could eventually go public (e.g.,
it would end up in the ".Rhistory" file). Instead of this, {gargle}'s encryption
function call an environment variable by its variable name and uses it to
encrypt the token.

To set the encryption key as an environment variable, it is required that it is
loaded to the R session on startup. This is usually done by setting it in a
".Renviron" file in the root folder of the project[^1].

[^1]: The ".Renviron" file, containing the environment variable with the secret
    key value needs, obviously, to be deployed along with the rest of the app
    code and ancillary files. Why this file is considered to be secure is
    something that I ignore.

### Procedure

1.  Create a ".Renviron" file in the root folder of the project

```{r create-renviron}
usethis::edit_r_environ(scope = "project")
```

2.  Write the encryption key to the ".Renviron" file with an object name that
    clearly identifies it, as `<KEY_VAR_NAME> = <secret_key>`, where
    `<KEY_VAR_NAME>` is the name of the object and `<secret_key>` is the secret
    key value (an alphanumeric string)

```{r write-envvar-key}
KEY_VAR_NAME <- "PID_PEC_IAD_KEY"
writeLines(paste(KEY_VAR_NAME, secret_key, sep = ' = '), con = ".Renviron")
```

3.  Restart the R session (from within Rstudio it can be done by calling)

```{r restart-R}
rstudioapi::restartSession(clean = FALSE)
```

4.  Verify that the key is created as an environment variable

```{r check-evvar-key}
#| eval: true

# NOTE: If the R session fails to restore, the following lines will restore the
#       missing objects and set up the options manually:
CACHE_PATH   <- ".secrets"
options(gargle_oauth_cache = CACHE_PATH)
gs_token     <- googlesheets4::gs4_token() # Select existing token if necessary
KEY_VAR_NAME <- "PID_PEC_IAD_KEY"

# Verify that the environment variable has the encryption key value
Sys.getenv(KEY_VAR_NAME)
```

## Encrypt the OAuth token

As explained before, we use this key to encrypt the OAuth token.

To do this, we will write the token as an encrypted R object, calling {gargle}'s
[`secret_write_rds()`](https://gargle.r-lib.org/reference/gargle_secret.html).
As explained before, this is done calling the encryption key by its environment
variable name, instead of its value. The function call will internally resolve
its value and use it to encrypt the token as an R object, and then save it as a
".rds" object.

Also, as the "decrypted" token will no longer be necessary, we will then delete
it to avoid its accidental exposure.

### Procedure

1.  Encrypt the token

```{r encrypt-token}
ENCRYPTED_TOKEN_FILENAME <- "encrypted-oauth-token.rds"
encrypted_token_filepath <- here::here(CACHE_PATH, ENCRYPTED_TOKEN_FILENAME)
gargle::secret_write_rds(
  gs_token,
  path = encrypted_token_filepath,
  key  = KEY_VAR_NAME
)
```

2.  Delete the decrypted token file

```{r delete-decrypted-token}
# TODO: Find the file by browsing to the ".secrets" folder, then substitute
#       "<token_filename>" with its name.
file.remove(here::here(CACHE_PATH, "<token_filename>"))
```

# Test non-interactive authorization {#test}

Using the encrypted token requires reading the encrypted file and decrypting it
to obtain the actual OAuth token object. This token is then used to grant
{googlesheets4} access to the Spreadsheets API.

To test that everything works, it is recommended to test writing to a Google
Spreadsheets file. In order to do that, we will follow these steps:

1.  Create a Google Spreadsheets file

2.  Read and decrypt the OAuth token from the encrypted file

3.  Grant access to {googlesheets4} using the decrypted token

4.  Attempt writing to the spreadsheet file

## Create a Google Spreadsheets file

The file must be created under **the Google account that was be used to generate
the OAuth token**. Alternatively, the file can be created with a different
account, but "edit access" must be granted to the account used to generate the
token. Bear in mind that this account is the "identity" that the Shiny app will
use to interact with the spreadsheet (and the Google API in general, for that
matter).

It is also recommended that, starting from cell A1, the spreadsheet file has the
structure of the dataset we want to save in it. This can be done simply by
typing in the variable names in the first row of the dataset.

For this minimal example, we will create a spreadsheet with the following two
variables:

```{r dataset-vars}
VAR_NAMES <- c("timestamp", "full_name", "valid")
```

### Procedure

1.  Go to [Google Spreadsheets](https://docs.google.com/spreadsheets). If
    necessary, log in with a valid Google account.

2.  Under "Create spreadsheet" click on "Blank spreadsheet".

3.  In the field "Untitled spreadsheet" type in a name, e.g., "Shiny_app_data"
    (without the double quotes).

4.  In cells A1 and A2, respectively, type in the variable names, i.e.,
    `timestamp`, `full_name` and `valid`. (For better readability, this row can
    be formatted as **boldface**.)

5.  In the browser, press `CTRL + L` (in Windows) to go to the address bar and
    then `CTRL + C` to copy the spreadsheet file URL, which should include
    everything up until the last "slash" ("/") character. In this example:

```{r gs-url}
# TODO: Change line 276 to use a different Google Spreadsheets file created
#       following these instructions, if necessary.
GS_URL <-
  "https://docs.google.com/spreadsheets/d/1-PWidqK7TncYK0W9AQd0J0L9GUpxL4SMzYmNg7LYn00/"
```

(Change line 86 to use a different file created following these instructions, if
necessary.)

6.  (OPTIONALLY) In [the Google Drive
    view](https://drive.google.com/drive/u/0/my-drive), find the spreadsheet
    file and [share it](https://support.google.com/drive/answer/2494822) with
    other Google account(s) (e.g. collaborators), **granting them "Edit"
    permission**.

## Read and decrypt the OAuth token from the encrypted file

This consists only of reading the OAuth token with the decryption function
`secret_read_rds()` from the {gargle} package, which is the "symmetric" to
`gargle::secret_write_rds()` we used earlier to encrypt the token. It takes in
similar arguments (except for the original token object, obviously) and returns
the decrypted token as an R object.

### Procedure

1.  Read and decrypt OAuth token

```{r read-encrypted-token}
gs_token <- gargle::secret_read_rds(
  path = encrypted_token_filepath,
  key  = KEY_VAR_NAME
)
```

## Grant access to {googlesheets4} using the decrypted token

This step consists only of granting access to the Google Spreadsheets API to
{googlesheets4}. We do this by calling authorization function
`googlesheets4::gs4_auth()` with the decrypted OAuth token as an argument.

### Procedure

1.  Authorize {googlesheets4} with the decrypted token

```{r auth-gs4}
googlesheets4::gs4_auth(token = gs_token)
```

## Attempt writing to the spreadsheet file

Assuming the authorization was successful, {googlesheets4} now has write access
to the Google Spreadsheets API. We will test this by trying to write lines to
the spreadsheet test file we created earlier.

To do that, we will first create a "link" object to the file (an object of the
class `googlesheets4_spreadheet`), and then use it to write to that file.
Although this step is not strictly necessary (we could access the file directly
through its URL anytime we want to write to it), this procedure appears
throughout the documentation so I assume it must be better for some reason
(probably faster).

To make sure the operation was successful, we use the current time in the
variable `timestamp`. Then, we can open the spreadsheet file and check that its
value in the last line of the file coincides with the time when the new data was
created (i.e., when `data.frame()` was called).

### Procedure

1.  Get the link object to the file

```{r get-file-link}
gs_link <- googlesheets4::gs4_get(GS_URL)
```

2.  Attempt writing a line to the file

```{r gs-write-line}
test_data <- data.frame(   # The timestamp value will indicate when this code
  lubridate::now("CET") |> #   block is run
    lubridate::stamp("2024-12-01 06:00", orders = "%Y%Om%d%H%M")(),
  "John Doe",
  FALSE
) |>
  setNames(VAR_NAMES)

googlesheets4::sheet_append(gs_link, test_data)
```

3.  Verify that the line has been written, navigating to `{r} GS_URL` and
    checking the last value of `timestamp`.

# Complete code for setting up a connection to a Google Spreadsheets file

Here I present a summed up version of the code needed to set up a connection to
a spreadsheet file. Note the file must have been already created, and populated
with the variable names in the first row, as per [the previous section](#test).

Then, in the app code it is just a matter of interacting with the spreadsheet
file with object `gs_link`.

```{r connect-gs-file}
# Read & decrypt OAuth token
gs_token <- gargle::secret_read_rds(
  path = encrypted_token_filepath,
  key  = KEY_VAR_NAME
)

# Grant access with OAuth token
googlesheets4::gs4_auth(token = gs_token)

# Get file link
gs_link <- googlesheets4::gs4_get(GS_URL)
```
