---
title: "Google Spreadsheets connection instructions"
format: html
editor: source
---

# Introduction

This document explains how to authorize and configure a connection to Google
Spreadsheets, so it can be used as a back-end to store the application data.

The procedure consists of setting up and encrypted [OAuth
token](https://oauth.net/) for granting the Shiny app access to Google
Spreadsheets. After setting it up, the token can be read and used in the app for
gaining access to the Spreadsheets API through [the {googlesheets4} R
package](https://googlesheets4.tidyverse.org/).

## Attributions

This procedure uses [the {gargle} package](https://gargle.r-lib.org/) (an R
interface for the Google APIs) to authenticate. Its description is based on the
following sources, both of them articles in the package website:

-   [Non-interactive
    auth](https://gargle.r-lib.org/articles/non-interactive-auth.html#project-level-oauth-cache),
    specially [section "Project-level OAuth
    cache"](https://gargle.r-lib.org/articles/non-interactive-auth.html#project-level-oauth-cache)

-   [Managing tokens
    securely](https://gargle.r-lib.org/articles/managing-tokens-securely.html),
    specially [section "Encrypt an R
    object"](https://gargle.r-lib.org/articles/managing-tokens-securely.html#encrypt-an-r-object)

## Setup summary

The setup process consists of the following steps:

1.  Create a Google Spreadsheet file

2.  Set up an OAuth token

3.  Create an encryption key

4.  Set the encryption key as an environment variable

5.  Encrypt the OAuth token

# Detailed procedure

## Create a Google Spreadsheet file

The file must be created under the Google account that will be used to generate
the OAuth token. Bear in mind that this account is the "identity" that the Shiny
app will use to interact with the spreadsheet (and the Google API in general,
for that matter).

It is also recommended that, starting from cell A1, the spreadsheet file has the
structure of the dataset we want to save in it. This can be done simply by
typing in the variable names in the first row of the dataset.

For this minimal example, we will create a spreadsheet with the following two
variables:

```{r dataset-vars}
VAR_NAMES <- c("full_name", "valid")
```

### Procedure

1.  Go to [Google Spreadsheets](https://docs.google.com/spreadsheets). If
    necessary, log in with a valid Google account.

2.  Under "Create spreadsheet" click on "Blank spreadsheet".

3.  In the field "Untitled spreadsheet" type in a name, e.g., "Shiny_app_data"
    (without the double quotes).

4.  In cells A1 and A2, respectively, type in the variable names, i.e.,
    `full_name` and `valid`. (For better readability, this row can be formatted
    as **boldface**.)

5.  In the browser, press `CTRL + L` (in Windows) to go to the address bar and
    then `CTRL + C` to copy the spreadsheet file URL, which should include
    everything up until the last "slash" ("/") character. In this example:

```{r gs-url}
GS_URL <-
  "https://docs.google.com/spreadsheets/d/1-PWidqK7TncYK0W9AQd0J0L9GUpxL4SMzYmNg7LYn00/"
```

(Change line 86 to use a different file created following these instructions, if
necessary.)

6.  (OPTIONALLY) In [the Google Drive
    view](https://drive.google.com/drive/u/0/my-drive), find the spreadsheet
    file and [share it](https://support.google.com/drive/answer/2494822) with
    other Google account(s) (e.g. collaborators), **granting them "Edit"
    permission**.

## Set up an OAuth token

This step will create an OAuth token for the Shiny app to use non-interactively,
which is necessary to deploy it on [shinyapps.io](https://www.shinyapps.io/)
without human intervention.

We will make use of the interactive capabilities of {gargle} to generate a token
through a browser and cache it in a secret directory (i.e. folder) local to this
project.

### Procedure

1.  Designate a project-specific cache directory to be used by {gargle}

```{r setup-cache-dir}
CACHE_PATH <- ".secrets"
options(gargle_oauth_cache = here::here(CACHE_PATH))
```

2.  Purposely trigger interactive authorization by {googlesheets4}

```{r interactive-auth}
googlesheets4::gs4_auth()
```

3.  In the browser tab opened by the call to `googlesheets4::gs4_auth()`, click
    on your account of choice to "authorize *Tidyverse API Packages*".

4.  In the next page, click on the button "Continue"

5.  Check the box "View, edit, create and delete all you spreadsheets in Google
    Spreadsheets

6.  Click on the button "Continue"

7.  If the authorization has been sucessfull, a page will show the following
    message:

```         
Authentication complete. Please close this page and return to R.
```

Close this window/tab and return to Rstudio. Now {googlesheets4} can read the
token from the designated cache:

```{r read-token}
gs_token <- googlesheets4::gs4_token()
```

{googlesheets4} will use this token as long as it finds it in the cache folder,
without the need for interactively authenticating again.

## Create an encryption key

After completing the previous step, you will notice that a ".secrets" folder has
been created in the project root folder, which contains a file with a name
similar to "<token_id>\_<user_name>@gmail.com", where `<token_id>` is a
hexadecimal number with several digits, and `<user_name>` is the user name of
the Google account used to "authorize *Tidyverse API Packages*".

This file is the token that will be used to grant the Shiny app access to the
spreadsheet file. However, as this file needs to be deployed to
[shinyapps.io](https://www.shinyapps.io/) along with the app code and associated
files, it needs to be encrypted to prevent it from the risk of being exposed
publicly. This could allow malicious agents accessing not only the spreadsheet
for this project, but all other spreadsheets this account has access to.

Instead, the token is encrypted with a symmetric key (i.e. a "password") that
will be used in the app to decrypt it.

### Procedure

1.  Generate a secret key using {gargle}

```{r gen-key}
secret_key <- gargle::secret_make_key()
```

## Set the encryption key as an environment variable

Using the encryption key requires that it is set as an environment variable.
This prevents the key from being used by calling its value directly, which could
result in logging its value in log files that could eventually go public (e.g.,
it would end up in the ".Rhistory" file). Instead of this, {gargle}'s encryption
function call an environment variable by its variable name and uses it to
encrypt the token.

To set the encryption key as an environment variable, it is required that it is
loaded to the R session on startup. This is usually done by setting it in a
".Renviron" file in the root folder of the project[^1].

[^1]: The ".Renviron" file, containing the environment variable with the secret
    key value needs, obviously, to be deployed along with the rest of the app
    code and ancillary files. Why this file is considered to be secure is
    something that I ignore.

### Procedure

1.  Create a ".Renviron" file in the root folder of the project

```{r create-renviron}
usethis::edit_r_environ(scope = "project")
```

2.  Write the encryption key to the ".Renviron" file with an object name that
    clearly identifies it, as `<KEY_VAR_NAME> = <secret_key>`, where
    `<KEY_VAR_NAME>` is the name of the object and `<secret_key>` is the secret
    key value (an alphanumeric string)

```{r write-envvar-key}
KEY_VAR_NAME <- "PID_PEC_IAD_KEY"
writeLines(paste(KEY_VAR_NAME, secret_key, sep = ' = '), con = ".Renviron")
```

3.  Restart the R session (from within Rstudio it can be done by calling)

```{r restart-R}
rstudioapi::restartSession(clean = FALSE)
```

4.  Verify that the key is created as an environment variable

```{r}
# NOTE: If the R session fails to restore, the following lines will restore the
#       missing objects and set up the options manually:
VAR_NAMES    <- c("full_name", "valid")
GS_URL       <-
  "https://docs.google.com/spreadsheets/d/1-PWidqK7TncYK0W9AQd0J0L9GUpxL4SMzYmNg7LYn00/"
CACHE_PATH   <- ".secrets"
options(gargle_oauth_cache = here::here(CACHE_PATH))
gs_token     <- googlesheets4::gs4_token() # Select existing token if necessary
KEY_VAR_NAME <- "PID_PEC_IAD_KEY"

# Verify that the environment variable has the encryption key value
Sys.getenv(KEY_VAR_NAME)
```

## Encrypt the OAuth token

As explained before, we use this key to encrypt the OAuth token.

To do this, we will write the token as an encrypted R object, calling {gargle}'s
[`secret_write_rds()`](https://gargle.r-lib.org/reference/gargle_secret.html).
As explained before, this is done calling the encryption key by its environment
variable name, instead of its value. The function call will internally resolve
its value and use it to encrypt the token as an R object, and then save it as a
".rds" object.

Also, as the "unencrypted" token will no longer be necessary, we will then
delete it to avoid it from being accidentally exposed.

### Procedure

1.  Encrypt the token

```{r encrypt-token}
ENCRYPTED_TOKEN_FILENAME <- "encrypted-oauth-token.rds"

gargle::secret_write_rds(
  gs_token,
  here::here(CACHE_PATH, ENCRYPTED_TOKEN_FILENAME),
  key = KEY_VAR_NAME
)
```

2.  Delete the unencrypted token file

```{r delete-unencrypted-token}
# TODO: Find the file by browsing to the ".secrets" folder, then substitute
#       "<token_filename>" with its name.
file.remove(here::here(CACHE_PATH, "<token_filename>"))
```
