---
title:               "PEC Exercise Correction"
author:              "Daniel Morillo"
format:
  html:
    df-print:        paged
    toc:             true
    embed-resources: true
editor:              source
execute: 
  output:            false
  warning:           false
params:
  test_mode:         false
---

```{r setup}
#| include: false

# Set working directory
knitr::opts_knit$set(root.dir = here::here())

# Read in parameters
test_mode <- params$test_mode
```

```{r libraries}
#| include: false
library(googlesheets4)
library(dplyr)
library(lubridate)
library(stringr)
library(tidyr)
library(purrr)
library(here)
library(tibble)
library(readr)
library(roperators)
```

```{r constants}
#| include: false

# Shared file system objects:
DATA_FOLDER <- "dat"


# Constant objects for processing the downloads dataset:
TEACHING_TEAM_EMAILS_FILENAME <- "teaching_team_emails.csv"

# Constant objects for processing the student responses dataset:

## File system objects:
DATA_RESPONSES_FILENAME <- "Respuestas actividad optativa de JAMOVI.csv"
RESPONSE_VARS_FILENAME  <- "student_response_variables.csv"

## Variable values:
RELATIONSHIP_VALUES <- c("Inversa", "No tienen relación", "Directa")
HELP_VALUES         <- c("1 - Nada", "2 - Un poco", "3 - Algo", "4 - Mucho")
USEFULNESS_VALUES   <- c(
  "1 - Nada útiles",
  "2 - Poco útiles",
  "3 - Algo útiles",
  "4 - Muy útiles"
)
CLARITY_VALUES      <- c(
  "1 - Nada claras",
  "2 - Poco claras",
  "3 - Algo claras",
  "4 - Muy claras"
)
PROFESSOR_EMAIL_DOMAIN <- "@psi.uned.es"
```

```{r sources}
#| include: false
source("R/data_storage.R",   encoding = 'UTF-8')
source("R/simulated_data.R", encoding = 'UTF-8')
source("R/hash_emails.R",    encoding = 'UTF-8')
```

# Introduction

This document retrieves the information for correcting the exercise of the
educational innovation project for the IAD course.

Using two source datasets, it computes:

-   The correct answers to the exercise for each student

-   Whether they have answered correctly to the exercise

-   Whether each student fulfills the assessment criteria

-   Which students must receive the course compensation for their participation

The source datasets are:

-   The log of users that have downloaded their personalized, random dataset,
    from the dataset generation Shiny app.

-   The OpenLMS output dataset with the exercise survey responses.

## Testing mode

Test mode allows to render the notebook **with all the responses**. Test mode
can be turned on/off by editing the value of the parameter `test_mode` to
`true`/`false` in the document "yaml" front matter (line 12). When test mode is
off, the responses from the teaching team are filtered out applying the
following criteria:

-   From the student responses dataset, responses that match the "teacher email
    domain" are directly filtered out.

-   From the downloads dataset, a set of teacher emails is loaded, their
    corresponding hashes are computed, and the matching ones are filtered out.

# Data loading & preprocessing

## Dataset of personalized data downloads

```{r config-downlads-data}
## File system objects:
teaching_team_emails_filepath <- here(
  DATA_FOLDER,
  TEACHING_TEAM_EMAILS_FILENAME
)
```

The downloads are first read from the Google Spreadsheets file connection. Prior
to reading the dataset, one must make sure that [the Gsheets API connection has
been successfully authorized](doc/gs_instructions.qmd).

```{r load-download-data}
data_downloads <- gs_file_link |>
  read_sheet(col_types = 'icc') |> # Read email hashes as integers
  mutate(across(timestamp, parse_timestamp)) # `timestamp` as POSIXct time

if (!test_mode) {

  teaching_team_emails <- teaching_team_emails_filepath |> read_csv()

  # Hash emails for filtering out the matching hashes: 
  teaching_team_hashes <- teaching_team_emails |>
    transmute(email_hash = email_address |> map_int(hash_emails))

  # Filter out the teaching team emails (i.e. the ones included in the dataset
  #   with the "teaching team emails":
  data_downloads <- data_downloads |>
    anti_join(teaching_team_hashes, by = "email_hash")
}
```

Then, the data is processed to extract, for each user:

-   The number of access and download attempts.

-   The timestamp (in the local time of `{r} LOCAL_TIMEZONE`) of the first and
    last access and download attempts.

```{r process-download-data}
data_downloads_out <- data_downloads |>
  arrange(timestamp) |>
  summarise(
    attempts = n(),
    first    = first(timestamp),
    last     = last(timestamp),
    .by      = c(email_hash, event) # Compute for each event type and email hash
  ) |>
  mutate(across(event, str_to_lower)) |> # Lowercase events for column names
  pivot_wider(names_from = event, values_from = attempts:last)
```

## Dataset of activity responses

```{r config-responses-data}
## File system objects:
data_responses_filepath <- here(DATA_FOLDER, DATA_RESPONSES_FILENAME)
response_vars_filepath  <- here(DATA_FOLDER, RESPONSE_VARS_FILENAME)

# Variable name objects:
response_vars <- response_vars_filepath |> read_csv(col_types = "c")
response_var_labels <- response_vars |> deframe()
```

The dataset is first downloaded from [the OpenLMS
course](https://agora.uned.es/course/view.php?id=7595). To download them, we
need to access the [page of
responses](https://agora.uned.es/mod/feedback/show_entries.php?id=468562)
("Respuestas") in the activity survey, and download them by clicking the button
"Descargar". The format used will be "csv" (the default one). The data are saved
to [the data folder](dat/), with its default name,
"`{r} DATA_RESPONSES_FILENAME`".

This dataset is then read and preprocessed

```{r load-responses-data}
data_student_responses <- data_responses_filepath |>
  read_csv() |>
  rename(!!!response_var_labels) |>
  select(student_name, email_address:extra_comments) |>
  mutate(
    email_hash = email_address |> map_int(hash_emails),
    date = date |>
      parse_date_time(orders = "%d%B%Y %H%M") |>
      force_tz(tzone = LOCAL_TIMEZONE),
    relationship = relationship |> factor(
      levels = RELATIONSHIP_VALUES, labels = -1:1
    ),
    across(starts_with("help"), ~ordered(., levels = HELP_VALUES)),
    across(usefulness_videos,   ~ordered(., levels = USEFULNESS_VALUES)),
    across(instructions,        ~ordered(., levels = CLARITY_VALUES)),
    nps = nps |> str_extract("^\\d*") |> as.integer()
  )

if (!test_mode) {
  
  # Filter out the teaching team emails (i.e. the ones that match the "professor
  #   domain").
  data_student_responses <- data_student_responses |>
    filter(email_address |> str_detect(PROFESSOR_EMAIL_DOMAIN, negate = TRUE))
}
```

# Output computation

## Correct responses for each student

For each student, the correct answers of the exercise are obtained.

```{r get-correct-responses}
data_correct_responses <- data_downloads_out |>
  pull(email_hash) |>
  set_names() |>
  map(get_user_responses) |>
  list_rbind(names_to = "email_hash") |>
  mutate(across(email_hash, as.integer)) # Turn names to integers for matching
```

## Student responses correction

For each student, it is computed whether they have answered correctly to each of
the items in the exercise.

```{r correct-student-responses}
data_response_corrections <- data_student_responses |>
  select(email_hash, student_name, intercept:relationship) |>
  left_join(
    data_correct_responses,
    by     = "email_hash",
    suffix = c("_student", "_correct")
  ) |>
  mutate(
    intercept    = map2_lgl(intercept_student, intercept_correct, `%~=%`),
    slope        = map2_lgl(slope_student,     slope_correct,     `%~=%`),
    relationship = relationship_student %==% relationship_correct,
  )
```

## Assessment criteria fulfillment

To receive the course compensation, each student must have

-   downloaded their personalized dataset, and

-   completed the response survey of the activity.

We merge both datasets to detect the students that have done both steps.
Additionally, we check whether they have responded after downloading the dataset
(at least for the first time).

```{r compute-assessment-criteria}
#| output: true
data_criteria <- data_student_responses |>
  select(email_address, student_name, email_hash, date) |>
  left_join(
    data_downloads_out |> select(email_hash, first_download),
    by = "email_hash"
  ) |>
  mutate(
    download_correct = !is.na(first_download),
    response_correct = !is.na(date),
    date_correct     = first_download < date,
    criteria_all     = download_correct & response_correct & date_correct
  )

data_criteria |> count(across(download_correct:criteria_all))
```

The number of "TRUE" cases are the number of students that fulfill the
compensation criteria; the "FALSE" cases are the ones who do not.

## Students to receive compensation for participation

### Composition of the compensation dataset

We need the student ID and name to generate the dataset with the specified
format, but we have the email instead. Therefore, we need to match the email to
those fields.

#### Template workbook generation

```{r student-template-path}
STUDENT_TEMPLATE_FILENAME <- "Introducción al Análisis de Datos (011037)-006001.xls"

STUDENT_TEMPLATE_FILEPATH <- here(DATA_FOLDER, STUDENT_TEMPLATE_FILENAME)
```

We need to create an Excel workbook with a specific format. A workbook with this
format can be downloaded automatically from the Grading Management System, with
the following procedure:

1.  Authenticate in the UNED Campus for Teachers:
    https://www.uned.es/universidad/campus/pdi/docencia.html

2.  Browse to the Grading Management System:
    https://calificaciones.uned.es/sgc/index.asp

3.  Introduce the last 6 digits of the course code in the field "CÓDIGO
    ASIGNATURA": 011037

4.  The current course (2024-2025) is selected by default, so click in "Aceptar"

5.  Click on "Carga a través de excel" -\> "Generar ficheros excel"

6.  Activate checkboxes "TODOS LOS APELLIDOS" and "TODOS LOS CENTROS"

7.  Select radio button "Fichero excel agrupando por Centro Asociado"

8.  Textbox "Descripción del fichero" cannot be empty, so fill in with "Notas
    Jamovi", or similar

9.  Click on "Generar Ficheros"

10. When the progress bar indicates "100%", click on "Descargar Ficheros"

11. Save generated file as `r STUDENT_TEMPLATE_FILEPATH`

#### Student data extraction from the template workbook

Student data are extracted from the template workbook, including the name, and
ID. The name has to be processed to delete accents and reorder the full name as
"name surname", instead of "surname, name".

```{r read-student-sheet-template}
student_template <- readxl::read_excel(STUDENT_TEMPLATE_FILEPATH)

students <- student_template |>
  tidyr::separate(...3, into = c("surname", "name"), remove = FALSE, sep = ', ') |>
  dplyr::mutate(
    name      = name |> stringi::stri_trans_general("Latin-ASCII"),
    full_name = paste(name, surname)
  ) |>
  dplyr::filter(!is.na(...3), ...3 != "Apellidos, nombre")
```

```{r read-student-sheet-template}
students <- students |>
  dplyr::left_join(data_criteria, by = c(full_name = "student_name"))
```

#### Participant data generation

```{r participant-data-path}
PARTICIPANT_DATA_FILENAME <- "courseid_7595_participants.csv"

PARTICIPANT_DATA_FILEPATH <- here(DATA_FOLDER, PARTICIPANT_DATA_FILENAME)
```

1.  Browse to the virtual course in Agora:
    https://agora.uned.es/course/view.php?id=7595

2.  Click on tab "Participantes"

3.  Scroll down to the end of the page

4.  Click on "Seleccione a todos los usuarios <N>"

5.  Click on "Elegir..."

6.  Click on "**Descargar datos de tabla como**, Valores separados por comas
    (.csv)"

7.  Save the generated file as `r PARTICIPANT_DATA_FILEPATH`

### Compensation dataset creation

The dataset of the students that qualify for the compensation is created and
exported.

```{r create-compensation-dataset}
DATA_COMPENSATION_FILENAME <- "compensation_data.csv"

data_compensation_filepath <- here(DATA_FOLDER, DATA_COMPENSATION_FILENAME)

data_compensations <- data_criteria |>
  filter(criteria_all) |>
  select(email_address)

data_compensations |> write_csv(file = data_compensation_filepath)
```

