---
title:               "PEC Exercise Correction"
author:              "Daniel Morillo"
format:
  html:
    df-print:        paged
    toc:             true
    embed-resources: true
editor:              source
execute: 
  output:            false
  warning:           false
params:
  test_mode:         false
---

```{r setup}
#| include: false

# Set working directory
knitr::opts_knit$set(root.dir = here::here())

# Read in parameters
test_mode <- params$test_mode
```

```{r libraries}
#| include: false
library(here)
library(dplyr)
library(googlesheets4)
library(readr)
library(purrr)
library(tidyr)
library(stringr)
library(roperators)
library(writexl)
```

```{r sources}
#| include: false
source("R/constants.R",              encoding = 'UTF-8')
source("R/data_storage.R",           encoding = 'UTF-8')
source("R/score_responses.R",        encoding = 'UTF-8')
source("R/read_student_responses.R", encoding = 'UTF-8')
```

```{r constants}
#| include: false

# Constant objects for processing the downloads dataset:

## Dataset with the emails of the teaching team (to filter them out in
##   "non-test" mode)
TEACHING_TEAM_EMAILS_FILENAME <- "teaching_team_emails.csv"
teaching_team_emails_filepath <- here(
  DATA_DIR,
  TEACHING_TEAM_EMAILS_FILENAME
)

## File system objects for processing the student responses dataset:
DATA_RESPONSES_FILENAME <- "Respuestas actividad optativa de JAMOVI.csv"
data_responses_filepath <- here(DATA_DIR, DATA_RESPONSES_FILENAME)

## File system objects for generating the response dataset:
GRADING_TEMPLATE_FILENAME    <- # (Course name and activity code may change)
  "Calificaciones-Introducción al Análisis de Datos-Calificaciones-756233.csv"
DATA_GRADING_FILENAME        <- "grading_data.csv"

grading_template_filepath  <- here(DATA_DIR, GRADING_TEMPLATE_FILENAME)
data_grading_filepath      <- here(DATA_DIR, DATA_GRADING_FILENAME)
```

# Introduction

This document retrieves the information for correcting the exercise of the
educational innovation project for the IAD course.

Using two source datasets, it computes:

-   The correct answers to the exercise for each student

-   Whether they have answered correctly to each item of the exercise

-   Whether each student fulfills the assessment criteria

-   The grading of each student in the activity

The source datasets are:

-   The log of users that have downloaded their personalized, random dataset,
    from the dataset generation Shiny app.

-   The OpenLMS output dataset with the exercise survey responses.

Then, it creates a dataset with the students that receive the course
compensation. This requires adapting the compensation data to a template
generated from the UNED grading system. The format of this template include the
student name and personal ID, but the OpenLMS data only has the email, so the
data needs to be matched by name to generate the compensation with the tempate
format.

## Pre-requirements

The Shiny app log file is directly accessed through the Google sheets API.
Therefore, both the API and the log file must be configured according to the
Shiny app instructions.

In order to this, before running this notebook, make sure to follow:

1.  The [“Google Spreadsheets connection
    instructions”](../doc/gs_instructions.qmd)

2.  The ["Configuration" section](../README.md#configuration) of [the repository
    README file](../README.md)

## Testing mode

Test mode allows to render the notebook **with all the responses**. Test mode
can be turned on/off by editing the value of the parameter `test_mode` to
`true`/`false` in the document "yaml" front matter (line 12). When test mode is
off, the responses from the teaching team are filtered out applying the
following criteria:

-   From the student responses dataset, responses that match the "teacher email
    domain" are directly filtered out.

-   From the downloads dataset, a set of teacher emails is loaded, their
    corresponding hashes are computed, and the matching ones are filtered out.

# Data loading & preprocessing

## Dataset of personalized data downloads

The downloads are first read from the Google Spreadsheets file connection. Prior
to reading the dataset, one must make sure that [the Gsheets API connection has
been successfully authorized](doc/gs_instructions.qmd).

```{r load-download-data}
data_downloads <- gs_file_link |>
  read_sheet(col_types = 'icc') |> # Read email hashes as integers
  mutate(across(timestamp, parse_timestamp)) # `timestamp` as POSIXct time

if (!test_mode) {

  teaching_team_emails <- teaching_team_emails_filepath |> read_csv(
    col_types = cols(email_address = col_character()),
  )

  # Hash emails for filtering out the matching hashes: 
  teaching_team_hashes <- teaching_team_emails |> mutate(
    !!EMAIL_HASH_VAR := email_address |> map_int(hash_emails),
    .keep = "none"
  )

  # Filter out the teaching team emails (i.e. the ones included in the dataset
  #   with the "teaching team emails":
  data_downloads <- data_downloads |>
    anti_join(teaching_team_hashes, by = EMAIL_HASH_VAR)
}
```

Then, the data is processed to extract, for each user:

-   The number of access and download attempts.

-   The timestamp (in the local time of `{r} LOCAL_TIMEZONE`) of the first and
    last access and download attempts.

```{r process-download-data}
data_downloads_out <- data_downloads |>
  arrange(timestamp) |>
  summarise(
    attempts = n(),
    first    = first(timestamp),
    last     = last(timestamp),
    .by      = c(!!EMAIL_HASH_VAR, event) # Compute for each event type and email hash
  ) |>
  mutate(across(event, str_to_lower)) |> # Lowercase events for column names
  pivot_wider(names_from = event, values_from = attempts:last)
```

## Dataset of activity responses

The dataset is first downloaded from [the OpenLMS
course](https://agora.uned.es/course/view.php?id=12075). To download them, we
need to access the [page of
responses](https://agora.uned.es/mod/feedback/show_entries.php?id=625026)
("Respuestas") in the activity survey, and download them by clicking the button
"Descargar". The format used will be "csv" (the default one). Save the dataset
to [the data folder](dat/) with its default name,
"`{r} DATA_RESPONSES_FILENAME`".

::: callout-note
If you want to define any exemptions by date for a specific set of items, run
the `pec-exercise-exemptions.qmd` notebook at this point.
:::

This dataset is then read and preprocessed

```{r load-responses-data}
data_student_responses <- data_responses_filepath |>
  read_student_responses(test = test_mode, correct_num = TRUE) |>
  select(email_address, date, !!EMAIL_HASH_VAR, starts_with("item_"))
```

# Output computation

## Correct responses for each student

For each student, the correct answers of the exercise are obtained.

```{r get-correct-responses}
data_correct_responses <- data_student_responses |>
  pull(!!EMAIL_HASH_VAR) |>
  set_names() |> # Helps matching each user's responses with its hash
  map(get_correct_responses) |> # Get correct responses for each student
  list_rbind(names_to = EMAIL_HASH_VAR) |>
  mutate(across(!!EMAIL_HASH_VAR, as.integer)) # For matching purposes
```

## Student responses correction

For each student, it is computed whether they have answered correctly to each of
the items in the exercise. Then, the total score is computed as the sum of
correct answers.

```{r correct-student-responses}
data_response_corrections <- data_student_responses |>
  score_student_responses(data_correct_responses) |>
  rowwise() |>
  mutate(
    grading_bonus = ends_with(SCORE_RESPONSE_SUFFIX) |>
      c_across() |>
      mean() *
      BONUS_MAX
  )
```

## Download dataset checks

To confirm that each student has used their personalized dataset to solve the
activity, we check whether the students who have responded to the activity have
downloaded the dataset, and whether they have downloaded it (at least the first
time) before responding to the activity.

```{r compute-assessment-criteria}
#| output: true
data_criteria <- data_student_responses |>
  select(!!EMAIL_HASH_VAR, date) |>
  left_join(
    data_downloads_out |> select(!!EMAIL_HASH_VAR, first_download),
    by = EMAIL_HASH_VAR
  ) |>
  mutate(
    data_downloaded  = !is.na(first_download),
    response_correct = !is.na(date),
    # Operator `roperators::%<~%` returns `FALSE` if one of the operands is `NA`
    date_correct     = first_download %<~% date,
    criteria_all     = data_downloaded & response_correct & date_correct
  )

data_criteria |> count(across(data_downloaded:criteria_all))
```

The number of "TRUE" cases are the number of students that have downloaded the
dataset and can be graded; the "FALSE" cases are the ones who have not.

## Students to receive scoring for participation

### Composition of the grading dataset

We need to fill in a grading sheet in a ".csv" file, with a specific format
created by OpenLMS. In order to this, we first download the template from
OpenLMS, and then fill in the grading of the scored students.

#### Template grading dataset download

To download the template, we first need to create a "Task" in OpenLMS. This will
be a "virtual" activity (invisible for the students) where the grading sheet can
be uploaded.

We have created [this activity in
OpenLMS](https://agora.uned.es/mod/assign/view.php?id=756233), with the name
"Calificaciones", and set its visibility as "No mostrado a los estudiantes". To
download the grading sheet template, follow these steps:

1.  Authenticate in the UNED Campus for Teachers (if necessary):
    https://www.uned.es/universidad/campus/pdi/docencia.html

2.  Browse to the [grading page of the
    activity](https://agora.uned.es/mod/assign/view.php?id=756233&action=grading).

3.  Click on button "Acciones".

4.  In the drop-down menu, click on "Descargar la hoja de calificaciones".

5.  Save the file with its default name "`{r} GRADING_TEMPLATE_FILENAME`" in
    [the data folder](dat/).

```{r read-student-grading_template}
grading_template <- read_csv(
  grading_template_filepath,
  col_types = cols(.default = col_character()) # Prevent parsing default values
)
```

#### Grading dataset creation

The dataset with the scoring data is matched to the criteria dataset and, for
the students that fulfill the criteria, the score is written to the template. To
do this, we match by email hash all the datasets. After matching, we verify that
the score of the students who do not fulfill the criteria have the expeceted
score (i.e. zero, as they have probably responded at random).

```{r generate-grading-data}
grading_data <- data_response_corrections |>
  ungroup() |>
  select(email_hash, grading_bonus) |>
  left_join(
    data_criteria |> select(email_hash, data_downloaded:criteria_all),
    by = EMAIL_HASH_VAR
  )

# Check scores against grading criteria:
grading_data |> count(across(grading_bonus | data_downloaded:criteria_all))

# Select necessary variables and cases from the grading dataset
grading_data <- grading_data |>
  filter(criteria_all) |>
  select(!!EMAIL_HASH_VAR, grading_bonus)

# Fill in the grading template with the computed scores
grading_data <- grading_template |>
  mutate(!!EMAIL_HASH_VAR := `Dirección de correo` |> map_int(hash_emails)) |>
  left_join(grading_data, by = EMAIL_HASH_VAR) |>
  mutate(Calificación = grading_bonus) |>
  select(-!!EMAIL_HASH_VAR, -grading_bonus)

# Discard the cases that have not been graded (i.e., `Calificación` is `NA`)
grading_data <- grading_data |>
  filter(!is.na(Calificación))
```

### Grading data exportation

The grading dataset is exported to file "`{r} data_grading_filepath`".

```{r create-compensation-dataset}
grading_data |> write_excel_csv(data_grading_filepath)
```

### Grading dataset upload

The grading dataset can now be uploaded to [the activity created in
OpenLMS](https://agora.uned.es/mod/assign/view.php?id=756233) previously. To
upload the data file with the grading sheet template, follow these steps:

1.  Authenticate in the UNED Campus for Teachers (if necessary):
    https://www.uned.es/universidad/campus/pdi/docencia.html

2.  Browse to the [grading page of the
    activity](https://agora.uned.es/mod/assign/view.php?id=756233&action=grading).

3.  Click on button "Acciones".

4.  In the drop-down menu, click on "Subir la hoja de calificaciones".

5.  In the "file input", drag and drop the file named
    "`{r} DATA_GRADING_FILENAME`" in [the data folder](dat/).

6.  Check box "Permitir la actualización de los registros que se han modificado
    recientemente en Moodle antes que en la hoja de cálculo."

7.  Click on button "Subir la hoja de calificaciones".

8.  In the confirmation screen, click on "Confirmar".

## Student responses

In order to implement the feedback app, we need to create a file with the
student responses, so they can compare them with the correct solutions. In order
to do this, we export the student responses along with the email hash, so that
we can publish these responses without exposing the students' data.

```{r create-responses-dataset}
data_student_responses_out <- data_student_responses |>
  select(!!EMAIL_HASH_VAR, starts_with("item_")) |>
  mutate(across(where(is.double), ~round(., 2))) # Round numeric responses for
                                                 #   writing them

data_student_responses_out |> write_excel_csv(file = student_responses_filepath)
```
