---
title:               "PEC Exercise Correction Overwrite"
author:              "Alejandro Martínez-Mingo"
format:
  html:
    df-print:        paged
    toc:             true
    embed-resources: true
editor:              source
execute: 
  output:            false
  warning:           false
---

```{r setup}
#| include: false

# Set working directory
knitr::opts_knit$set(root.dir = here::here())

```

```{r libraries}
#| include: false
library(here)
library(dplyr)
library(readr)
library(purrr)
library(tidyr)
library(stringr)
library(roperators)
# library(stringi)
# library(tibble)
# library(readxl)
library(writexl)
```

```{r sources}
#| include: false
source("R/constants.R",              encoding = 'UTF-8')
source("R/score_responses.R",        encoding = 'UTF-8')
source("R/read_student_responses.R", encoding = 'UTF-8')
```

```{r constants}
#| include: false

# Constant objects for processing the downloads dataset:

## File system objects for processing the student responses dataset:
DATA_RESPONSES_FILENAME <- "Respuestas actividad optativa de JAMOVI.csv"
data_responses_filepath <- here(DATA_DIR, DATA_RESPONSES_FILENAME)

# Dates (local) that define the special-case window (inclusive)
SPECIAL_CASES_DATE_FROM <- as.Date("2025-11-03")
SPECIAL_CASES_DATE_TO   <- as.Date("2025-11-17")

# Items to overwrite
ITEMS_TO_OVERWRITE <- c("item_8", "item_9", "item_10")
```

# Introduction

This document overwrites the responses of the students in
`Respuestas actividad optativa de JAMOVI.csv` returning the same file, that
should be used in the `pec-exercise-correction.qmd` in the case of need.

# Data processing

We just need to import the dataset with the students responses to the activity

## Dataset of activity responses

The dataset is first downloaded from [the OpenLMS
course](https://agora.uned.es/course/view.php?id=12075). To download them, we
need to access the [page of
responses](https://agora.uned.es/mod/feedback/show_entries.php?id=625026)
("Respuestas") in the activity survey, and download them by clicking the button
"Descargar". The format used will be "csv" (the default one). Save the dataset
to [the data folder](dat/) with its default name,
"`{r} DATA_RESPONSES_FILENAME`".

This dataset is then read and preprocessed

```{r load-responses-data}
data_student_responses <- data_responses_filepath |>
  read_student_responses(test = F, correct_num = TRUE) |>
  select(email_address, date, !!EMAIL_HASH_VAR, starts_with("item_"))
```

Select special cases by submission date (day-level, inclusive range)

```{r select-special-cases}
data_student_responses_to_overwrite <- data_student_responses |>
  dplyr::filter(!is.na(date)) |>
  dplyr::mutate(date_day = as.Date(date, tz = LOCAL_TIMEZONE)) |>
  dplyr::filter(dplyr::between(date_day, SPECIAL_CASES_DATE_FROM, SPECIAL_CASES_DATE_TO)) |>
  dplyr::select(-date_day)
```

Compute the hash of all students to overwrite

```{r select-special-hashes}
special_cases_hashes <- data_student_responses_to_overwrite |>
  dplyr::transmute(!!EMAIL_HASH_VAR := as.integer(.data[[EMAIL_HASH_VAR]])) |>
  dplyr::distinct()
```

# Output

## Compute correct questions and overwrite

For each student, the correct answers of the exercise are obtained.

```{r get-correct-responses}
data_correct_responses <- special_cases_hashes |>
  pull(!!EMAIL_HASH_VAR) |>
  set_names() |>                  # Helps matching each user's responses with its hash
  map(get_correct_responses) |>    # Get correct responses for each selected student
  list_rbind(names_to = EMAIL_HASH_VAR) |>
  mutate(across(!!EMAIL_HASH_VAR, as.integer)) # For matching purposes
```

```{r refine-special-cases-by-reversed-model}
# Helper funtion: compute correct answers when DV/IV are swapped (only items 8-10)
compute_correct_responses_reversed <- function(data) {

  INTERCEPT_TERM <- "(Intercept)"

  # Swapped model vs the official one in compute_correct_responses()
  criterion_name <- SIM_VAR_NAMES["var_2"] |> glue::backtick()
  predictor_name <- SIM_VAR_NAMES["var_1"] |> glue::backtick()

  # Same transformation used in compute_correct_responses() (var_2 as integer)
  data <- data |>
    dplyr::mutate(
      `Satisfaccion vital` = `Satisfaccion vital` |>
        as.character() |>
        as.integer()
    )

  model        <- glue::glue("{criterion_name} ~ {predictor_name}")
  fitted_model <- stats::lm(formula = model, data = data)
  coefficients <- broom::tidy(fitted_model)

  out <- data |> dplyr::summarise(
    item_8  = coefficients |>
      dplyr::filter(term == INTERCEPT_TERM) |>
      dplyr::pull(estimate) |>
      round(N_DECIMALS),
    item_9  = coefficients |>
      dplyr::filter(term == predictor_name) |>
      dplyr::pull(estimate) |>
      round(N_DECIMALS),
    item_10 = item_9 |> sign() |> as.character()
  )

  out <- out |> dplyr::mutate(
    item_10 = item_10 |>
      readr::parse_factor(levels = ITEM_10_VALUES) |>
      forcats::fct_recode(!!!ITEM_10_VALUES)
  )
  
  out
}

get_correct_responses_reversed <- function(hash) {
  simulate_data(hash) |> compute_correct_responses_reversed()
}

# Compute reversed keys for the same date-window candidates
data_correct_responses_rev <- special_cases_hashes |>
  dplyr::pull(!!EMAIL_HASH_VAR) |>
  set_names() |>
  purrr::map(get_correct_responses_reversed) |>
  purrr::list_rbind(names_to = EMAIL_HASH_VAR) |>
  dplyr::mutate(across(!!EMAIL_HASH_VAR, as.integer)) |>
  dplyr::select(!!EMAIL_HASH_VAR, dplyr::all_of(ITEMS_TO_OVERWRITE)) |>
  dplyr::rename_with(~ paste0(.x, "_rev"), dplyr::all_of(ITEMS_TO_OVERWRITE))

data_correct_responses_fwd <- data_correct_responses |>
  dplyr::select(!!EMAIL_HASH_VAR, dplyr::all_of(ITEMS_TO_OVERWRITE)) |>
  dplyr::rename_with(~ paste0(.x, "_fwd"), dplyr::all_of(ITEMS_TO_OVERWRITE))

# Compare student answers vs forward/reversed keys to detect "reversed solvers"
candidates <- data_student_responses_to_overwrite |>
  dplyr::select(!!EMAIL_HASH_VAR, dplyr::all_of(ITEMS_TO_OVERWRITE)) |>
  dplyr::distinct(.data[[EMAIL_HASH_VAR]], .keep_all = TRUE) |>
  dplyr::left_join(data_correct_responses_fwd, by = EMAIL_HASH_VAR) |>
  dplyr::left_join(data_correct_responses_rev, by = EMAIL_HASH_VAR)

for (it in ITEMS_TO_OVERWRITE) {
  candidates[[paste0(it, "_match_fwd")]] <- purrr::map2_lgl(
    candidates[[it]], candidates[[paste0(it, "_fwd")]],
    roperators::`%~=%`
  ) |> dplyr::coalesce(FALSE)

  candidates[[paste0(it, "_match_rev")]] <- purrr::map2_lgl(
    candidates[[it]], candidates[[paste0(it, "_rev")]],
    roperators::`%~=%`
  ) |> dplyr::coalesce(FALSE)
}

# Apply exemption only if they match the reversed model
candidates <- candidates |>
  dplyr::mutate(
    rev_ok = item_8_match_rev | item_9_match_rev,
    apply_exemption = rev_ok
  )

# Refine the special cases to overwrite
special_cases_hashes <- candidates |>
  dplyr::filter(apply_exemption) |>
  dplyr::transmute(!!EMAIL_HASH_VAR := as.integer(.data[[EMAIL_HASH_VAR]])) |>
  dplyr::distinct()

# Keep only the forward answer key for the truly exempted cases (used to overwrite to "officially correct")
data_correct_responses <- data_correct_responses |>
  dplyr::semi_join(special_cases_hashes, by = EMAIL_HASH_VAR)



```

Next, we calculate the correct answers for the special cases:

```{r calculate-correct-responses-special-cases}
# Build overrides: correct answers for ONLY the special-case students and ONLY the affected items
special_overrides <- data_correct_responses |>
  dplyr::select(dplyr::all_of(c(EMAIL_HASH_VAR, ITEMS_TO_OVERWRITE))) |>
  dplyr::distinct(.data[[EMAIL_HASH_VAR]], .keep_all = TRUE) |>
  dplyr::rename_with(
    ~ paste0(.x, "_override"),
    dplyr::all_of(ITEMS_TO_OVERWRITE)
  )
```

In the next chunk we overwrite the special cases. I found some problems since
some types were coerced during the overwritting process, so we need to check the
type of every item we want to overwrite to ensure it is the same as in the
original data.

```{r overwrite-special-cases}
# Join override columns into the main responses (do NOT change types here)
data_student_responses_overwritten <- data_student_responses |>
  dplyr::left_join(special_overrides, by = EMAIL_HASH_VAR)

# Overwrite item-by-item, preserving the original column type
for (item in ITEMS_TO_OVERWRITE) {
  override_col <- paste0(item, "_override")
  if (!override_col %in% names(data_student_responses_overwritten)) next

  # Cast override values to the same type as the target column to avoid vctrs upcasting.
  target <- data_student_responses_overwritten[[item]]
  ov     <- data_student_responses_overwritten[[override_col]]

  if (is.integer(target)) {
    ov <- as.integer(ov)
  } else if (is.numeric(target)) {
    ov <- as.numeric(ov)
  } else if (is.logical(target)) {
    ov <- as.logical(ov)
  } else {
    ov <- as.character(ov)
  }

  idx <- !is.na(ov)
  data_student_responses_overwritten[[item]][idx] <- ov[idx]
}

# Drop helper override columns
data_student_responses_overwritten <- data_student_responses_overwritten |>
  dplyr::select(-dplyr::any_of(paste0(ITEMS_TO_OVERWRITE, "_override")))
```

Now, `data_student_responses_overwritten` should have the new overwritten
responses in the selected items for the selected students and can be used
directly in the `pec-exercise-correction.qmd` notebook, but we assume that the
original file should be overwritten. For doing this in the raw format we need to
load the raw data first.

```{r load-responses-data-raw}
data_student_responses_raw <- readr::read_csv(
  data_responses_filepath,
  show_col_types = FALSE,
  col_types = readr::cols(.default = readr::col_character())
)
```

We input the new data in the original format.

```{r overwrite-raw-and-export}
# Map standard vars (item_*) to raw column names using the same mapping as read_student_responses()
raw_email_col <- unname(response_vars_labels["email_address"])
raw_item_cols <- unname(response_vars_labels[ITEMS_TO_OVERWRITE])
names(raw_item_cols) <- ITEMS_TO_OVERWRITE

# Create a function to format numeric values to match the raw CSV style (decimal comma)
fmt_num_es <- function(x, digits = N_DECIMALS) {
  x <- suppressWarnings(as.numeric(x))
  ifelse(
    is.na(x),
    NA_character_,
    format(round(x, digits), nsmall = digits, decimal.mark = ",",
           scientific = FALSE, trim = TRUE)
  )
}

# Build the updates table
updates_raw <- data_student_responses_overwritten |>
  dplyr::semi_join(special_cases_hashes, by = EMAIL_HASH_VAR) |>
  dplyr::select(email_hash, dplyr::all_of(ITEMS_TO_OVERWRITE)) |>
  dplyr::distinct(email_hash, .keep_all = TRUE) |>
  dplyr::mutate(
    dplyr::across(
      dplyr::all_of(ITEMS_TO_OVERWRITE),
      \(x) dplyr::case_when(
        is.numeric(x)  ~ fmt_num_es(x),
        is.integer(x)  ~ fmt_num_es(x),
        is.factor(x)   ~ as.character(x),
        is.logical(x)  ~ as.character(x),
        TRUE           ~ as.character(x)
      )
    )
  ) |>
  dplyr::rename(!!!stats::setNames(ITEMS_TO_OVERWRITE, raw_item_cols))

# Compute hashes for RAW rows, locate rows, and overwrite cells
raw_hash <- purrr::map_int(
  data_student_responses_raw[[raw_email_col]],
  \(x) if (is.na(x) || x == "") NA_integer_ else hash_emails(x)
)
idx <- match(updates_raw$email_hash, raw_hash)
ok  <- !is.na(idx)
for (raw_col in unname(raw_item_cols)) {
  data_student_responses_raw[[raw_col]][idx[ok]] <- updates_raw[[raw_col]][ok]
}
```

Finally we export the new datafile overwritting the original one

```{r data-export}
# Export using the original filename (raw format)
readr::write_excel_csv(data_student_responses_raw, data_responses_filepath)
```

Additionally, we also export the emails of the students whose responses have
been changed to notify them about the correction.

```{r}
overrides_emails <- data_student_responses |>
  semi_join(special_overrides, by = "email_hash") |>
  select(email_address)

readr::write_excel_csv(overrides_emails, here(DATA_DIR, "overrides_emails.csv"))
```

## Check dataset overwriting

The overwritten dataset is tested with the reading function, to check that this
notebook can be run and inserted safely in the "correction process".

```{r test-exported-responses}
data_student_responses <- data_responses_filepath |>
  read_student_responses(correct_num = TRUE)
```
