---
title:               "PEC Exercise Correction"
author:              "Daniel Morillo"
format:
  html:
    df-print:        paged
    toc:             true
    embed-resources: true
editor:              source
execute: 
  output:            false
  warning:           false
---

```{r setup}
#| include: false

# Set working directory
knitr::opts_knit$set(root.dir = here::here())
```

```{r libraries}
#| include: false
library(here)
library(dplyr)
library(googlesheets4)
library(readr)
library(purrr)
library(tidyr)
library(stringr)
library(roperators)
# library(stringi)
# library(tibble)
# library(readxl)
library(writexl)
```

```{r sources}
#| include: false
source("R/constants.R",              encoding = 'UTF-8')
source("R/data_storage.R",           encoding = 'UTF-8')
source("R/score_responses.R",        encoding = 'UTF-8')
source("R/read_student_responses.R", encoding = 'UTF-8')
```

```{r constants}
#| include: false

# Constant objects for processing the downloads dataset:

## Dataset with the emails of the teaching team (to filter them out in
##   "non-test" mode)
TEACHING_TEAM_EMAILS_FILENAME <- "teaching_team_emails.csv"
teaching_team_emails_filepath <- here(
  DATA_DIR,
  TEACHING_TEAM_EMAILS_FILENAME
)

## File system objects for processing the student responses dataset:
DATA_RESPONSES_FILENAME <- "Respuestas actividad optativa de JAMOVI.csv"
data_responses_filepath <- here(DATA_DIR, DATA_RESPONSES_FILENAME)

## File system objects for generating the response dataset:
GRADING_TEMPLATE_FILENAME    <- # (Course name and activity code may change)
  "Calificaciones-Introducción al Análisis de Datos-Calificaciones-756233.csv"
DATA_GRADING_FILENAME        <- "grading_data.csv"
STUDENT_RESPONSES_FILENAME   <- "student_responses.csv"

grading_template_filepath  <- here(DATA_DIR, GRADING_TEMPLATE_FILENAME)
data_grading_filepath      <- here(DATA_DIR, DATA_GRADING_FILENAME)
```

# Introduction

This document retrieves the information for correcting the exercise of the
educational innovation project for the IAD course.

Using two source datasets, it computes:

-   The correct answers to the exercise for each student

-   Whether they have answered correctly to each item of the exercise

-   Whether each student fulfills the assessment criteria

-   The grading of each student in the activity

The source datasets are:

-   The log of users that have downloaded their personalized, random dataset,
    from the dataset generation Shiny app.

-   The OpenLMS output dataset with the exercise survey responses.

Then, it creates a dataset with the students that receive the course
compensation. This requires adapting the compensation data to a template
generated from the UNED grading system. The format of this template include the
student name and personal ID, but the OpenLMS data only has the email, so the
data needs to be matched by name to generate the compensation with the tempate
format.

# Data loading & preprocessing

## Dataset of activity responses

The dataset is first downloaded from [the OpenLMS
course](https://agora.uned.es/course/view.php?id=12075). To download them, we
need to access the [page of
responses](https://agora.uned.es/mod/feedback/show_entries.php?id=625026)
("Respuestas") in the activity survey, and download them by clicking the button
"Descargar". The format used will be "csv" (the default one). Save the dataset
to [the data folder](dat/) with its default name,
"`{r} DATA_RESPONSES_FILENAME`".

This dataset is then read and preprocessed

```{r load-responses-data}
data_student_responses <- data_responses_filepath |>
  read_student_responses(test = test_mode, correct_num = TRUE) |>
  select(email_address, date, !!EMAIL_HASH_VAR, starts_with("item_"))
```

# Output computation

## Student responses correction

For each student, it is computed whether they have answered correctly to each of
the items in the exercise. Then, the total score is computed as the sum of
correct answers.

```{r}
compute_correct_responses <- function(data) {
  ## Constant objects: ----

  # Linear regression model objects:
  INTERCEPT_TERM <- "(Intercept)"
  criterion_name <- SIM_VAR_NAMES['var_1'] |> glue::backtick()
  predictor_name <- SIM_VAR_NAMES['var_2'] |> glue::backtick()


  ## Main: ----

  # Transform `var_2` to integer to use it a "linear term" in the regression
  data <- data |> dplyr::mutate(`Satisfaccion vital` = `Satisfaccion vital` |>
                                  as.character() |>
                                  as.integer())

  # Fit model and extract coefficients:
  model       <- glue::glue("{criterion_name} ~ {predictor_name}") # Formula
  fitted_model <- data         |> lm(formula = model)
  coefficients <- fitted_model |> broom::tidy()

  # Compute responses:

  computed_responses <- data |> dplyr::summarize(
    item_1 = table(`Satisfaccion vital`)[ITEM_1_LS_CAT],
    item_2 = median(`Horas sueno promedio`),
    item_3 = sd(`Horas sueno promedio`) |> round(N_DECIMALS),
    item_4 = quantile(`Horas sueno promedio`, .34),
    item_5 = boxplot.stats(`Horas sueno promedio`)$out |>
      length() |>
      as.logical() |>
      as.character(),
    item_6 = cor(`Satisfaccion vital`, `Horas sueno promedio`) |>
      round(N_DECIMALS),
    item_7 = item_6 |> sign() |> as.character(),
    item_8 = coefficients |>
      dplyr::filter(term == INTERCEPT_TERM) |>
      dplyr::pull(estimate) |>
      round(N_DECIMALS),
    item_9 = coefficients |>
      dplyr::filter(term == predictor_name) |>
      dplyr::pull(estimate) |>
      round(N_DECIMALS),
    item_10 = item_9 |> sign() |> as.character(),
  )

  # Capture warning when factor levels are missing in the data (they will!)
  suppressWarnings(
    computed_responses <- computed_responses |> dplyr::mutate(
      item_5  = item_5 |>
        readr::parse_factor(levels = ITEM_5_VALUES) |>
        forcats::fct_recode(!!!ITEM_5_VALUES),
      item_7  = item_7 |>
        readr::parse_factor(levels = ITEM_7_VALUES) |>
        forcats::fct_recode(!!!ITEM_7_VALUES),
      item_10 = item_10 |>
        readr::parse_factor(levels = ITEM_10_VALUES) |>
        forcats::fct_recode(!!!ITEM_10_VALUES)
    )
  )

  computed_responses
}
```

## Download dataset checks

To confirm that each student has used their personalized dataset to solve the
activity, we check whether the students who have responded to the activity have
downloaded the dataset, and whether they have downloaded it (at least the first
time) before responding to the activity.

```{r compute-assessment-criteria}
#| output: true
data_criteria <- data_student_responses |>
  select(!!EMAIL_HASH_VAR, date) |>
  left_join(
    data_downloads_out |> select(!!EMAIL_HASH_VAR, first_download),
    by = EMAIL_HASH_VAR
  ) |>
  mutate(
    data_downloaded  = !is.na(first_download),
    response_correct = !is.na(date),
    # Operator `roperators::%<~%` returns `FALSE` if one of the operands is `NA`
    date_correct     = first_download %<~% date,
    criteria_all     = data_downloaded & response_correct & date_correct
  )

data_criteria |> count(across(data_downloaded:criteria_all))
```

The number of "TRUE" cases are the number of students that have downloaded the
dataset and can be graded; the "FALSE" cases are the ones who have not.

## Student responses

In order to implement the feedback app, we need to create a file with the
student responses, so they can compare them with the correct solutions. In order
to do this, we export the student responses along with the email hash, so that
we can publish these responses without exposing the students' data.

```{r create-responses-dataset}
# We use the root folder to avoid accidental exposure of the "dat" folder
student_responses_filepath <- here(STUDENT_RESPONSES_FILENAME)

data_student_responses_out <- data_student_responses |>
  select(!!EMAIL_HASH_VAR, starts_with("item_")) |>
  mutate(across(where(is.double), ~round(., 2))) # Round numeric responses for
                                                 #   writing them

data_student_responses_out |> write_csv(file = student_responses_filepath)
```
