---
title:               "PEC Exercise Correction"
author:              "Daniel Morillo"
format:
  html:
    df-print:        paged
    toc:             true
    embed-resources: true
editor:              source
execute: 
  output:            false
  warning:           false
params:
  test_mode:         false
---

```{r setup}
#| include: false

# Set working directory
knitr::opts_knit$set(root.dir = here::here())

# Read in parameters
test_mode <- params$test_mode
```

```{r libraries}
#| include: false
library(googlesheets4)
library(dplyr)
library(lubridate)
library(stringr)
library(tidyr)
library(purrr)
library(here)
library(tibble)
library(readr)
library(roperators)
library(readxl)
library(stringi)
library(writexl)
```

```{r constants}
#| include: false

# Shared file system objects:
DATA_FOLDER <- "dat"


# Constant objects for processing the downloads dataset:
TEACHING_TEAM_EMAILS_FILENAME <- "teaching_team_emails.csv"

# Constant objects for processing the student responses dataset:

## File system objects:
DATA_RESPONSES_FILENAME <- "Respuestas actividad optativa de JAMOVI.csv"
RESPONSE_VARS_FILENAME  <- "student_response_variables.csv"

## Variable values:
RELATIONSHIP_VALUES <- c("Inversa", "No tienen relación", "Directa")
HELP_VALUES         <- c("1 - Nada", "2 - Un poco", "3 - Algo", "4 - Mucho")
USEFULNESS_VALUES   <- c(
  "1 - Nada útiles",
  "2 - Poco útiles",
  "3 - Algo útiles",
  "4 - Muy útiles"
)
CLARITY_VALUES      <- c(
  "1 - Nada claras",
  "2 - Poco claras",
  "3 - Algo claras",
  "4 - Muy claras"
)
PROFESSOR_EMAIL_DOMAIN <- "@psi.uned.es"

# Constant objects for generating the response dataset:

## File system objects:

STUDENT_TEMPLATE_FILENAME    <- # NOTE: Last code (006121) may change
  "Introducción al Análisis de Datos (011037)-006401.xls"
COURSE_PARTICIPANTS_FILENAME <- "courseid_7595_participants.csv"
DATA_COMPENSATION_FILENAME   <- "compensation_data.xls"
STUDENT_RESPONSES_FILENAME   <- "student_responses.csv"
```

```{r sources}
#| include: false
source("R/data_storage.R",   encoding = 'UTF-8')
source("R/simulated_data.R", encoding = 'UTF-8')
source("R/hash_emails.R",    encoding = 'UTF-8')
```

# Introduction

This document retrieves the information for correcting the exercise of the
educational innovation project for the IAD course.

Using two source datasets, it computes:

-   The correct answers to the exercise for each student

-   Whether they have answered correctly to each item of the exercise

-   Whether each student fulfills the assessment criteria

-   The grading of each student in the activity

The source datasets are:

-   The log of users that have downloaded their personalized, random dataset,
    from the dataset generation Shiny app.

-   The OpenLMS output dataset with the exercise survey responses.

Then, it creates a dataset with the students that receive the course
compensation. This requires adapting the compensation data to a template
generated from the UNED grading system. The format of this template include the
student name and personal ID, but the OpenLMS data only has the email, so the
data needs to be matched by name to generate the compensation with the tempate
format.

## Testing mode

Test mode allows to render the notebook **with all the responses**. Test mode
can be turned on/off by editing the value of the parameter `test_mode` to
`true`/`false` in the document "yaml" front matter (line 12). When test mode is
off, the responses from the teaching team are filtered out applying the
following criteria:

-   From the student responses dataset, responses that match the "teacher email
    domain" are directly filtered out.

-   From the downloads dataset, a set of teacher emails is loaded, their
    corresponding hashes are computed, and the matching ones are filtered out.

# Data loading & preprocessing

## Dataset of personalized data downloads

```{r config-downlads-data}
## File system objects:
teaching_team_emails_filepath <- here(
  DATA_FOLDER,
  TEACHING_TEAM_EMAILS_FILENAME
)
```

The downloads are first read from the Google Spreadsheets file connection. Prior
to reading the dataset, one must make sure that [the Gsheets API connection has
been successfully authorized](doc/gs_instructions.qmd).

```{r load-download-data}
data_downloads <- gs_file_link |>
  read_sheet(col_types = 'icc') |> # Read email hashes as integers
  mutate(across(timestamp, parse_timestamp)) # `timestamp` as POSIXct time

if (!test_mode) {

  teaching_team_emails <- teaching_team_emails_filepath |> read_csv()

  # Hash emails for filtering out the matching hashes: 
  teaching_team_hashes <- teaching_team_emails |>
    transmute(email_hash = email_address |> map_int(hash_emails))

  # Filter out the teaching team emails (i.e. the ones included in the dataset
  #   with the "teaching team emails":
  data_downloads <- data_downloads |>
    anti_join(teaching_team_hashes, by = "email_hash")
}
```

Then, the data is processed to extract, for each user:

-   The number of access and download attempts.

-   The timestamp (in the local time of `{r} LOCAL_TIMEZONE`) of the first and
    last access and download attempts.

```{r process-download-data}
data_downloads_out <- data_downloads |>
  arrange(timestamp) |>
  summarise(
    attempts = n(),
    first    = first(timestamp),
    last     = last(timestamp),
    .by      = c(email_hash, event) # Compute for each event type and email hash
  ) |>
  mutate(across(event, str_to_lower)) |> # Lowercase events for column names
  pivot_wider(names_from = event, values_from = attempts:last)
```

## Dataset of activity responses

```{r config-responses-data}
## File system objects:
data_responses_filepath <- here(DATA_FOLDER, DATA_RESPONSES_FILENAME)
response_vars_filepath  <- here(DATA_FOLDER, RESPONSE_VARS_FILENAME)

# Variable name objects:
response_vars       <- response_vars_filepath |> read_csv(col_types = "c")
response_var_labels <- response_vars |> deframe()
```

The dataset is first downloaded from [the OpenLMS
course](https://agora.uned.es/course/view.php?id=12075). To download them, we
need to access the [page of
responses](https://agora.uned.es/mod/feedback/show_entries.php?id=625026)
("Respuestas") in the activity survey, and download them by clicking the button
"Descargar". The format used will be "csv" (the default one). Save the dataset
to [the data folder](dat/) with its default name,
"`{r} DATA_RESPONSES_FILENAME`".

This dataset is then read and preprocessed

```{r load-responses-data}
data_student_responses <- data_responses_filepath |>
  read_csv() |>
  rename(!!!response_var_labels) |>
  select(student_name, email_address:extra_comments) |>
  mutate(
    email_hash = email_address |> map_int(hash_emails),
    date = date |>
      parse_date_time(orders = "%d%B%Y %H%M") |>
      force_tz(tzone = LOCAL_TIMEZONE),
    ## TODO: Format responses
    # relationship = relationship |> factor(
    #   levels = RELATIONSHIP_VALUES, labels = -1:1
    # ),
    # across(starts_with("help"), ~ordered(., levels = HELP_VALUES)),
    # across(usefulness_videos,   ~ordered(., levels = USEFULNESS_VALUES)),
    # across(instructions,        ~ordered(., levels = CLARITY_VALUES)),
    nps = nps |> str_extract("^\\d*") |> as.integer()
  )

if (!test_mode) {
  
  # Filter out the teaching team emails (i.e. the ones that match the "professor
  #   domain").
  data_student_responses <- data_student_responses |>
    filter(email_address |> str_detect(PROFESSOR_EMAIL_DOMAIN, negate = TRUE))
}
```

# Output computation

## Correct responses for each student

For each student, the correct answers of the exercise are obtained.

```{r get-correct-responses}
data_correct_responses <- data_downloads_out |>
  pull(email_hash) |>
  set_names() |>
  map(get_user_responses) |> # Get correct responses for each student
  list_rbind(names_to = "email_hash") |>
  mutate(across(email_hash, as.integer)) # Turn names to integers for matching
```

## Student responses correction

For each student, it is computed whether they have answered correctly to each of
the items in the exercise.

```{r correct-student-responses}
data_response_corrections <- data_student_responses |>
  select(email_hash, student_name, intercept:relationship) |>
  left_join(
    data_correct_responses,
    by     = "email_hash",
    suffix = c("_student", "_correct")
  ) |>
  mutate(
    intercept    = map2_lgl(intercept_student, intercept_correct, `%~=%`),
    slope        = map2_lgl(slope_student,     slope_correct,     `%~=%`),
    relationship = relationship_student %==% relationship_correct,
  )
```

## Assessment criteria fulfillment

To receive the course compensation, each student must have

-   downloaded their personalized dataset, and

-   completed the response survey of the activity.

We merge both datasets to detect the students that have done both steps.
Additionally, we check whether they have responded after downloading the dataset
(at least for the first time).

```{r compute-assessment-criteria}
#| output: true
data_criteria <- data_student_responses |>
  select(student_name, email_hash, date) |>
  left_join(
    data_downloads_out |> select(email_hash, first_download),
    by = "email_hash"
  ) |>
  mutate(
    download_correct = !is.na(first_download),
    response_correct = !is.na(date),
    date_correct     = first_download < date,
    criteria_all     = download_correct & response_correct & date_correct
  )

data_criteria |> count(across(download_correct:criteria_all))
```

The number of "TRUE" cases are the number of students that fulfill the
compensation criteria; the "FALSE" cases are the ones who do not.

## Students to receive compensation for participation

### Composition of the compensation dataset

We need the student's personal ID (DNI / NIE) and name to generate the dataset
with the specified format, but we have the email instead. Therefore, we need to
match the email to those fields.

#### Template workbook generation

We need to create an Excel workbook with a specific format. A workbook with this
format can be downloaded automatically from the Grading Management System, with
the following procedure:

1.  Authenticate in the UNED Campus for Teachers:
    https://www.uned.es/universidad/campus/pdi/docencia.html

2.  Browse to the Grading Management System:
    https://calificaciones.uned.es/sgc/index.asp

3.  Introduce the last 6 digits of the course code in the field "CÓDIGO
    ASIGNATURA": `011037`, then click on "Aceptar"

4.  The current course (2024-2025) is selected by default, so click on "Aceptar"

5.  Click on "Carga a través de excel" -\> "Generar ficheros excel"

6.  Activate checkboxes "TODOS LOS APELLIDOS" and "TODOS LOS CENTROS"

7.  Select radio button "Fichero excel agrupando por Centro Asociado"

8.  Textbox "Descripción del fichero" cannot be empty, so fill in with "Notas
    Jamovi", or similar

9.  Click on "Generar Ficheros"

10. When the progress bar indicates "100%", click on "Descargar Ficheros"

```{r student-template-path}
student_template_filepath <- here(DATA_FOLDER, STUDENT_TEMPLATE_FILENAME)
```

11. Save generated file as "`{r} student_template_filepath`"

12. Read in the template file

```{r read-student-sheet-template}
student_template <- read_excel(
  student_template_filepath,
  range = cell_limits(c(1, 1), c(NA, NA)), # Read the full sheet
  col_names = c(
    "empty_1", "empty_2",            # Empty cells (respect in template)
    "center", "id", "surname_name",  # Student identification data
    "empty_6", "empty_7", "empty_8", # Empty cells (respect in template)
    "grade"                          # Grades (to fill in the grading values)
  )
)
```

#### Student data extraction from the template workbook

Student data are extracted from the template workbook, including the name, and
ID. The name has to be processed to delete accents and reorder the full name as
"name surname", instead of "surname, name".

```{r read-student-name-id}
student_data <- student_template |>
  select(id, surname_name) |>
  dplyr::filter(!is.na(id), id != "D.N.I.") |>
  separate(
    surname_name,
    into = c("surname", "name"),
    remove = FALSE,
    sep = ', '
  ) |>
  mutate(surname = surname |> str_trim()) |>
  unite(student_name, name, surname, sep = ' ') |>
  mutate(student_name = student_name |> stri_trans_char("ÁÉÍÓÚ", "AEIOU"))
```

#### Compensation dataset creation

The dataset with the student data is matched to the criteria dataset. To do
this, we need to process the criteria dataset, filtering only the cases that
fulfill the compensation criteria, and matching by name. An additional column
with the compensation is added to the resulting dataset.

```{r generate-compensation-data}
data_compensation <- data_criteria |>
  filter(criteria_all) |>
  select(student_name) |>
  mutate(student_name = student_name |> stri_trans_char("ÁÉÍÓÚ", "AEIOU"))

student_compensation_data <- student_data |>
  semi_join(data_compensation, by = "student_name") |>
  add_column(grade = '0,2') |>
  select(-student_name)
```

### Data integrity checks

The following checks are performed:

-   Whether there are repeated names in the student dataset after applying the
    compensation

```{r check-duplicate-students}
#| output: true
dup_students_check <- student_compensation_data |>
  group_by(surname_name) |>
  filter(n() != 1)

n_dup_students <- dup_students_check |> nrow()

dup_students_check
```

There are `{r} n_dup_students` students who receive compensation that have the
same name as another student who also receives compensation.

-   Whether there are repeated names in the compensation results dataset:

```{r check-duplicates-compensation}
#| output: true
dup_data_compensation <- data_compensation |>
  group_by(student_name) |>
  filter(n() != 1)

n_dup_compensation <- dup_data_compensation |> nrow()

dup_data_compensation
```

There are `{r} n_dup_compensation` students who qualify for compensation that
have the same name as another student who also qualifies for compensation.

(If the previous two checks do not match, that means that the compensation of
one student is being applied to two or more students.)

-   Whether there are unmatched cases in the compensation data:

```{r unmatched-compensation-students}
#| output: true
unmatched_compensation <- data_compensation |>
  anti_join(student_data, by = join_by(student_name))

n_unmatched <- unmatched_compensation |> nrow()

unmatched_compensation
```

There are `{r} n_unmatched` students who completed the activity and qualify for
compensation, but the complete name does not match its equivalent in the student
template (due to missing compound names, different ways of writing them, or
missing second surnames) or does not exist.

### Compensation dataset exportation

```{r data-compensation-path}
data_compensation_filepath <- here(DATA_FOLDER, DATA_COMPENSATION_FILENAME)
```

The template with the student data is updated with the compensation data. Then,
the resulting dataset is exported to file "`{r} data_compensation_filepath`".

```{r create-compensation-dataset}
student_template <- student_template |>
  rows_update(student_compensation_data, by = c("id", "surname_name"))

student_template |> write_xlsx(
  data_compensation_filepath |> paste0('x'),
  col_names = FALSE,
  format_headers = FALSE
)
```

## Student responses

In order to implement the feedback app, we need to create a file with the
student responses, so they can compare them with the correct solutions. In order
to do this, we export the student responses along with the email hash, so that
we can publish these responses without exposing the students' data.

```{r create-responses-dataset}
student_responses_filepath <- here(STUDENT_RESPONSES_FILENAME)

data_student_responses_out <- data_student_responses |>
  select(email_hash, intercept, slope, relationship, interpretation)

data_student_responses_out |> write_csv(file = student_responses_filepath)
```
