---
title: "PEC Exercise Correction"
author: "Daniel Morillo"
format: html
editor: source
format:
  html:
    df-print: paged
execute: 
  output:     false
  warning:    false
---

```{r setup}
#| include: false

# Set working directory
knitr::opts_knit$set(root.dir = here::here())
```

```{r libraries}
#| include: false
library(googlesheets4)
library(dplyr)
library(lubridate)
library(stringr)
library(tidyr)
library(purrr)
library(here)
library(tibble)
library(readr)
library(roperators)
```

```{r constants}
#| include: false

# Shared file system objects:
DATA_FOLDER <- "dat"


# Constant objects for processing the downloads dataset:
TEACHING_TEAM_EMAILS_FILENAME <- "teaching_team_emails.csv"

# Constant objects for processing the student responses dataset:

## File system objects:
DATA_RESPONSES_FILENAME <- "Respuestas actividad optativa de JAMOVI.csv"
RESPONSE_VARS_FILENAME  <- "student_response_variables.csv"

## Variable values:
RELATIONSHIP_VALUES <- c("Inversa", "No tienen relación", "Directa")
HELP_VALUES         <- c("1 - Nada", "2 - Un poco", "3 - Algo", "4 - Mucho")
USEFULNESS_VALUES   <- c(
  "1 - Nada útiles",
  "2 - Poco útiles",
  "3 - Algo útiles",
  "4 - Muy útiles"
)
CLARITY_VALUES      <- c(
  "1 - Nada claras",
  "2 - Poco claras",
  "3 - Algo claras",
  "4 - Muy claras"
)
PROFESSOR_EMAIL_DOMAIN <- "@psi.uned.es"
```

```{r sources}
#| include: false
source("R/data_storage.R",   encoding = 'UTF-8')
source("R/simulated_data.R", encoding = 'UTF-8')
source("R/hash_emails.R",    encoding = 'UTF-8')
```

# Introduction

This document retrieves the information for correcting the exercise of the
educational innovation project for the IAD course.

Using two source datasets, it computes:

-   The correct answers to the exercise for each student

-   Whether they have answered correctly to the exercise

-   Whether each student fulfills the assessment criteria

-   Which students must receive the course compensation for their participation

The source datasets are:

-   The log of users that have downloaded their personalized, random dataset,
    from the dataset generation Shiny app.

-   The OpenLMS output dataset with the exercise survey responses.

# Data loading & preprocessing

## Dataset of personalized data downloads

The downloads are first read from the Google Spreadsheets file connection. Prior
to reading the dataset, one must make sure that [the Gsheets API connection has
been successfully authorized](doc/gs_instructions.qmd).

```{r load-download-data}
data_downloads <- gs_file_link |>
  read_sheet(col_types = 'icc') |> # Read email hashes as integers
  mutate(across(timestamp, parse_timestamp)) # `timestamp` as POSIXct time
```

Then, the data is processed to extract, for each user:

-   The number of access and download attempts.

-   The timestamp (in the local time of `{r} LOCAL_TIMEZONE`) of the first and
    last access and download attempts.

```{r process-download-data}
data_downloads_out <- data_downloads |>
  arrange(timestamp) |>
  summarise(
    attempts = n(),
    first    = first(timestamp),
    last     = last(timestamp),
    .by      = c(email_hash, event) # Compute for each event type and email hash
  ) |>
  mutate(across(event, str_to_lower)) |> # Lowercase events for column names
  pivot_wider(names_from = event, values_from = attempts:last)
```

## Dataset of activity responses

```{r config-responses-data}
## File system objects:
data_responses_filepath <- here(DATA_FOLDER, DATA_RESPONSES_FILENAME)
response_vars_filepath  <- here(DATA_FOLDER, RESPONSE_VARS_FILENAME)

# Variable name objects:
response_vars <- response_vars_filepath |> read_csv(col_types = "c")
response_var_labels <- response_vars |> deframe()
```

The dataset is first downloaded from [the OpenLMS
course](https://agora.uned.es/course/view.php?id=7595). To download them, we
need to access the page of responses ("Respuestas") in the activity survey, and
download them by clicking the button "Descargar". The format used will be "csv"
(the default one). The data are saved to [the data folder](dat/), with its
default name, "`{r} DATA_RESPONSES_FILENAME`".

This dataset is then read and preprocessed

```{r load-responses-data}
data_student_responses <- data_responses_filepath |>
  read_csv() |>
  rename(!!!response_var_labels) |>
  select(email_address:extra_comments) |>
  mutate(
    email_hash = email_address |> map_int(hash_emails),
    date = date |>
      parse_date_time(orders = "%d%B%Y %H%M") |>
      force_tz(tzone = LOCAL_TIMEZONE),
    relationship = relationship |> factor(
      levels = RELATIONSHIP_VALUES, labels = -1:1
    ),
    across(starts_with("help"), ordered, levels = HELP_VALUES),
    across(usefulness_videos,   ordered, levels = USEFULNESS_VALUES),
    across(instructions,        ordered, levels = CLARITY_VALUES),
    nps = nps |> str_extract("^\\d*") |> as.integer()
  )
```

# Output computation

## Correct responses for each student

For each student, the correct answers of the exercise are obtained.

```{r get-correct-responses}
data_correct_responses <- data_downloads_out |>
  pull(email_hash) |>
  set_names() |>
  map(get_user_responses) |>
  list_rbind(names_to = "email_hash") |>
  mutate(across(email_hash, as.integer)) # Turn names to integers for matching
```

## Student responses correction

For each student, it is computed whether they have answered correctly to each of
the items in the exercise.

```{r correct-student-responses}
data_response_corrections <- data_student_responses |>
  select(email_hash, intercept:relationship) |>
  left_join(
    data_correct_responses,
    by     = "email_hash",
    suffix = c("_student", "_correct")
  ) |>
  mutate(
    intercept    = map2_lgl(intercept_student, intercept_correct, `%~=%`),
    slope        = map2_lgl(slope_student,     slope_correct,     `%~=%`),
    relationship = relationship_student %==% relationship_correct,
  )
```

## Assessment criteria fulfillment

To receive the course compensation, each student must have

-   downloaded their personalized dataset, and

-   completed the response survey of the activity.

We merge both datasets to detect the students that have done both steps.
Additionally, we check whether they have responded after downloading the dataset
(at least for the first time).

```{r compute-assessment-criteria}
#| output: true
data_criteria <- data_student_responses |>
  select(email_address, email_hash, date) |>
  full_join(
    data_downloads_out |> select(email_hash, first_download),
    by = "email_hash"
  ) |>
  mutate(
    download_correct = !is.na(first_download),
    response_correct = !is.na(date),
    date_correct     = first_download < date,
    criteria_all     = download_correct & response_correct & date_correct
  )

data_criteria |> count(criteria_all)
```

## Students to receive compensation for participation

The dataset of the students that qualify for the compensation is created and
exported.

```{r create-compensation-dataset}
DATA_COMPENSATION_FILENAME <- "compensation_data.csv"

data_compensation_filepath <- here(DATA_FOLDER, DATA_COMPENSATION_FILENAME)

data_compensations <- data_criteria |>
  filter(criteria_all) |>
  select(email_address)

data_compensations |> write_csv(file = data_compensation_filepath)
```
