---
title:               "PEC Exercise Correction Overwrite"
author:              "Alejandro Martínez-Mingo"
format:
  html:
    df-print:        paged
    toc:             true
    embed-resources: true
editor:              source
execute: 
  output:            false
  warning:           false
---

```{r setup}
#| include: false

# Set working directory
knitr::opts_knit$set(root.dir = here::here())

```

```{r libraries}
#| include: false
library(here)
library(dplyr)
library(googlesheets4)
library(readr)
library(purrr)
library(tidyr)
library(stringr)
library(roperators)
# library(stringi)
# library(tibble)
# library(readxl)
library(writexl)
```

```{r sources}
#| include: false
source("R/constants.R",              encoding = 'UTF-8')
source("R/data_storage.R",           encoding = 'UTF-8')
source("R/score_responses.R",        encoding = 'UTF-8')
source("R/read_student_responses.R", encoding = 'UTF-8')
```

```{r constants}
#| include: false

# Constant objects for processing the downloads dataset:



## File system objects for processing the student responses dataset:
DATA_RESPONSES_FILENAME <- "Respuestas actividad optativa de JAMOVI.csv"
data_responses_filepath <- here(DATA_DIR, DATA_RESPONSES_FILENAME)
DATA_RESPONSES_OVERWRITE_FILENAME <- "Respuestas actividad optativa de JAMOVI to overwrite.csv"
data_responses_overwrite_filepath <- here(DATA_DIR, DATA_RESPONSES_OVERWRITE_FILENAME)

# Items to overwrite
ITEMS_TO_OVERWRITE <- c("item_8", "item_9", "item_10")

## File system objects for generating the response dataset:
DATA_RESPONSES_OVERWRITTEN_FILENAME   <- "Respuestas actividad optativa de JAMOVI overwritten.csv"
data_responses_overwritten_filepath  <- here(DATA_DIR, DATA_RESPONSES_OVERWRITTEN_FILENAME)
```

# Introduction

This document overwrites the responses of the students in
`Respuestas actividad optativa de JAMOVI to overwrite.csv` in the
`Respuestas actividad optativa de JAMOVI.csv` returning the file
`Respuestas actividad optativa de JAMOVI overwriten.csv` that should be used in
the `pec-exercise-correction.qmd` in the case of need.

# Data loading & preprocessing

## Dataset of personalized data downloads

The downloads are first read from the Google Spreadsheets file connection. Prior
to reading the dataset, one must make sure that [the Gsheets API connection has
been successfully authorized](doc/gs_instructions.qmd).

```{r load-download-data}
data_downloads <- gs_file_link |>
  read_sheet(col_types = 'icc') |> # Read email hashes as integers
  mutate(across(timestamp, parse_timestamp)) # `timestamp` as POSIXct time
```

Then, the data is processed to extract, for each user:

-   The number of access and download attempts.

-   The timestamp (in the local time of `{r} LOCAL_TIMEZONE`) of the first and
    last access and download attempts.

```{r process-download-data}
data_downloads_out <- data_downloads |>
  arrange(timestamp) |>
  summarise(
    attempts = n(),
    first    = first(timestamp),
    last     = last(timestamp),
    .by      = c(!!EMAIL_HASH_VAR, event) # Compute for each event type and email hash
  ) |>
  mutate(across(event, str_to_lower)) |> # Lowercase events for column names
  pivot_wider(names_from = event, values_from = attempts:last)
```

## Dataset of activity responses

The dataset is first downloaded from [the OpenLMS
course](https://agora.uned.es/course/view.php?id=12075). To download them, we
need to access the [page of
responses](https://agora.uned.es/mod/feedback/show_entries.php?id=625026)
("Respuestas") in the activity survey, and download them by clicking the button
"Descargar". The format used will be "csv" (the default one). Save the dataset
to [the data folder](dat/) with its default name,
"`{r} DATA_RESPONSES_FILENAME`".

This dataset is then read and preprocessed

```{r load-responses-data}
data_student_responses <- data_responses_filepath |>
  read_student_responses(test = F, correct_num = TRUE) |>
  select(email_address, date, !!EMAIL_HASH_VAR, starts_with("item_"))
```

To overwrite responses we need to load the special cases

```{r load-overwrite-hashes}
#| include: false

# Read the overwrite file and extract the hashes to modify
overwrite_raw <- readr::read_csv(
  data_responses_overwrite_filepath,
  show_col_types = FALSE
)

# Obtain the special cases hash
special_cases_hashes <- data_responses_overwrite_filepath |>
    read_student_responses(test = F, correct_num = TRUE) |>
    dplyr::transmute(!!EMAIL_HASH_VAR := as.integer(.data[[EMAIL_HASH_VAR]])) 
```

Now we can check how many cases should be overwritten by comparing the hashes
with the original students responses table (this should work also with a list of
students emails).

```{r test-n-cases-found}
h_students <- unique(na.omit(data_student_responses$email_hash))
h_special  <- unique(na.omit(special_cases_hashes$email_hash))

overlap_hashes <- intersect(h_students, h_special)

cat("nº students to overwrite: ", length(h_special), "\nnº students to overwrite found in data: ", length(overlap_hashes))
```

If we find all students in the data file, we can select those to change their
responses.

```{r test-cases-to-overwrite}
data_students_responses_to_overwrite <- data_student_responses %>%
  semi_join(special_cases_hashes, by = "email_hash")

cat("The number of rows (",nrow(data_students_responses_to_overwrite), ") of students responses to overwrite should be =", length(h_special))
```

# Output

## Compute correct questions and overwrite

For each student, the correct answers of the exercise are obtained.

```{r get-correct-responses}
data_correct_responses <- data_downloads_out |>
  # Filter out the ones without responses to make it more efficient:
  semi_join(data_student_responses, by = EMAIL_HASH_VAR) |>
  pull(!!EMAIL_HASH_VAR) |>
  set_names() |> # Helps matching each user's responses with its hash
  map(get_correct_responses) |> # Get correct responses for each student
  list_rbind(names_to = EMAIL_HASH_VAR) |>
  mutate(across(!!EMAIL_HASH_VAR, as.integer)) # For matching purposes
```

Next, we calculate the correct answers for the special cases:

```{r calculate-correct-responses-special-cases}
# Build overrides: correct answers for ONLY the special-case students and ONLY the affected items
special_overrides <- data_correct_responses |>
  dplyr::semi_join(special_cases_hashes, by = EMAIL_HASH_VAR) |>
  dplyr::select(dplyr::all_of(c(EMAIL_HASH_VAR, ITEMS_TO_OVERWRITE))) |>
  dplyr::distinct(.data[[EMAIL_HASH_VAR]], .keep_all = TRUE) |>
  dplyr::rename_with(
    ~ paste0(.x, "_override"),
    dplyr::all_of(ITEMS_TO_OVERWRITE)
  )
```

In the next chunk we overwrite the special cases

```{r overwrite-special-cases}
#| include: false

# Join override columns into the main responses (do NOT change types here)
data_student_responses_overwritten <- data_student_responses |>
  dplyr::left_join(special_overrides, by = EMAIL_HASH_VAR)

# Overwrite item-by-item, preserving the original column type
for (item in ITEMS_TO_OVERWRITE) {
  override_col <- paste0(item, "_override")
  if (!override_col %in% names(data_student_responses_overwritten)) next

  # Cast override values to the same type as the target column to avoid vctrs upcasting.
  # (Key point: never coerce the *target* column to character.)
  target <- data_student_responses_overwritten[[item]]
  ov     <- data_student_responses_overwritten[[override_col]]

  if (is.integer(target)) {
    ov <- as.integer(ov)
  } else if (is.numeric(target)) {
    ov <- as.numeric(ov)
  } else if (is.logical(target)) {
    ov <- as.logical(ov)
  } else {
    ov <- as.character(ov)
  }

  idx <- !is.na(ov)
  data_student_responses_overwritten[[item]][idx] <- ov[idx]
}

# Drop helper override columns
data_student_responses_overwritten <- data_student_responses_overwritten |>
  dplyr::select(-dplyr::any_of(paste0(ITEMS_TO_OVERWRITE, "_override")))
```

We audit the cases we overwrote.

```{r audit-overwrite, message=FALSE}
eh    <- EMAIL_HASH_VAR
items <- ITEMS_TO_OVERWRITE

# Coverage: are all special-case hashes present in both datasets?
h_special <- unique(na.omit(special_cases_hashes[[eh]]))
missing_in_original <- setdiff(h_special, unique(na.omit(data_student_responses[[eh]])))
missing_in_new      <- setdiff(h_special, unique(na.omit(data_student_responses_overwritten[[eh]])))

cat("special_cases_expected:", length(h_special),
    "\nmissing_in_original:", length(missing_in_original),
    "\nmissing_in_overwritten:", length(missing_in_new), "\n")

# Build comparison table: ORIGINAL vs OVERWRITTEN + OVERRIDES (special cases only)
orig_sc <- data_student_responses |>
  dplyr::semi_join(special_cases_hashes, by = eh) |>
  dplyr::select(dplyr::all_of(c(eh, items))) |>
  dplyr::distinct(.data[[eh]], .keep_all = TRUE) |>
  dplyr::rename_with(~ paste0(.x, "_orig"), dplyr::all_of(items))

new_sc <- data_student_responses_overwritten |>
  dplyr::semi_join(special_cases_hashes, by = eh) |>
  dplyr::select(dplyr::all_of(c(eh, items))) |>
  dplyr::distinct(.data[[eh]], .keep_all = TRUE) |>
  dplyr::rename_with(~ paste0(.x, "_new"), dplyr::all_of(items))

over_sc <- special_overrides |>
  dplyr::semi_join(special_cases_hashes, by = eh) |>
  dplyr::select(dplyr::all_of(c(eh, paste0(items, "_override")))) |>
  dplyr::distinct(.data[[eh]], .keep_all = TRUE)

audit <- orig_sc |>
  dplyr::full_join(new_sc, by = eh) |>
  dplyr::left_join(over_sc, by = eh)

# Flags: NEW matches OVERRIDE (when override exists) + whether ORIGINAL changed to NEW
for (it in items) {
  o <- paste0(it, "_override"); n <- paste0(it, "_new"); g <- paste0(it, "_orig")
  audit[[paste0(it, "_matches_override")]] <- ifelse(is.na(audit[[o]]), NA, audit[[n]] == audit[[o]])
  audit[[paste0(it, "_changed")]]          <- audit[[g]] != audit[[n]]
}

match_cols   <- paste0(items, "_matches_override")
changed_cols <- paste0(items, "_changed")

# Summary: correctness vs override
match_summary <- dplyr::tibble(
  item    = items,
  n_tested = sapply(items, \(it) sum(!is.na(audit[[paste0(it, "_override")]]))),
  n_match  = sapply(items, \(it) sum(audit[[paste0(it, "_matches_override")]] %in% TRUE,  na.rm = TRUE)),
  n_fail   = sapply(items, \(it) sum(audit[[paste0(it, "_matches_override")]] %in% FALSE, na.rm = TRUE))
)
print(match_summary)

# Summary: how many changed from original to overwritten
chg_mat <- sapply(items, \(it) audit[[paste0(it, "_changed")]] %in% TRUE)
change_summary <- dplyr::tibble(
  item = items,
  n_changed = colSums(chg_mat)
) |>
  dplyr::bind_rows(dplyr::tibble(item = "any",  n_changed = sum(rowSums(chg_mat) > 0)),
                   dplyr::tibble(item = "all3", n_changed = sum(rowSums(chg_mat) == length(items))))
print(change_summary)

# Show failing overwrites (should be empty if everything is correct)
fails <- audit |>
  dplyr::filter(dplyr::if_any(dplyr::all_of(match_cols), ~ .x %in% FALSE)) |>
  dplyr::select(dplyr::all_of(eh), dplyr::ends_with("_new"), dplyr::ends_with("_override"), dplyr::all_of(match_cols))

cat("n_fail_rows:", nrow(fails), "\n")
if (nrow(fails) > 0) print(fails)
```

Now, `data_student_responses` should have the new overwritten responses in items
8 to 10 for the selected students. The last step will be to write the new file
with the overwritten responses. For doing this in the raw format we need to load
the raw data first.

```{r load-responses-data-raw}
#| include: false

data_student_responses_raw <- readr::read_csv(
  data_responses_filepath,
  show_col_types = FALSE,
  col_types = readr::cols(.default = readr::col_character())
)
```

We input the new data in the original format.

```{r overwrite-raw-and-export}
#| include: false

# Locate raw columns (email + questions 8–10)
pick_one_col <- function(df, pattern) {
  hits <- names(df) |>
    purrr::keep(~ stringr::str_detect(.x, stringr::regex(pattern, ignore_case = TRUE)))
  stopifnot(length(hits) == 1L)
  hits[[1]]
}

EMAIL_ADDRESS_RAW_VAR <- pick_one_col(
  data_student_responses_raw,
  "^Dirección de correo$|Direcci[oó]n de correo"
)

ITEM_8_RAW_VAR  <- pick_one_col(data_student_responses_raw, "^Realice una regresi[oó]n lineal")
ITEM_9_RAW_VAR  <- pick_one_col(data_student_responses_raw, "^Continuando con la regresi[oó]n lineal")
ITEM_10_RAW_VAR <- pick_one_col(
  data_student_responses_raw,
  "¿C[oó]mo interpretar[ií]a el resultado de la regresi[oó]n|interpretar[ií]a.*regresi[oó]n"
)

# Format numeric values like the original raw CSV (decimal comma + fixed decimals)
fmt_num_es <- function(x, digits = N_DECIMALS) {
  x <- suppressWarnings(as.numeric(x))
  ifelse(
    is.na(x),
    NA_character_,
    format(round(x, digits), nsmall = digits, decimal.mark = ",", scientific = FALSE, trim = TRUE)
  )
}

# Build raw-format overrides from `special_overrides` (already computed above)
special_overrides_raw <- special_overrides |>
  dplyr::transmute(
    email_hash = .data[[EMAIL_HASH_VAR]],
    item_8_raw_override  = fmt_num_es(item_8_override),
    item_9_raw_override  = fmt_num_es(item_9_override),
    item_10_raw_override = as.character(item_10_override)
  )

# Keep the exact original raw column order
raw_cols <- names(data_student_responses_raw)

# Compute hash in raw, join overrides, and overwrite ONLY when an override exists
data_student_responses_raw_overwritten <- data_student_responses_raw |>
  dplyr::mutate(
    email_hash = purrr::map_int(
      .data[[EMAIL_ADDRESS_RAW_VAR]],
      \(x) if (is.na(x) || x == "") NA_integer_ else hash_emails(x)
    )
  ) |>
  dplyr::left_join(special_overrides_raw, by = "email_hash") |>
  dplyr::mutate(
    !!ITEM_8_RAW_VAR  := dplyr::coalesce(item_8_raw_override,  .data[[ITEM_8_RAW_VAR]]),
    !!ITEM_9_RAW_VAR  := dplyr::coalesce(item_9_raw_override,  .data[[ITEM_9_RAW_VAR]]),
    !!ITEM_10_RAW_VAR := dplyr::coalesce(item_10_raw_override, .data[[ITEM_10_RAW_VAR]])
  ) |>
  dplyr::select(dplyr::all_of(raw_cols)) # Preserve EXACT raw columns and order
```

Finally we export the new datafile

```{r data-export}
# Export using the requested filename (raw format)
readr::write_excel_csv(data_student_responses_raw_overwritten, data_responses_overwritten_filepath)
```
